---
title: "Benchmarking Against `{SuperLearner}` and `{sl3}`"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


It would be nice to compare `{SuperLearner}` `{sl3}` and `{nadir}` on some small,
medium, and large-ish datasets to see how they compare in timing results. 
For now, we'll start with some smaller ones to see how long it takes to get 
`super_learner()` and `cv_super_learner()` (and their equivalents) to run on these. 

For these different sizes, I'd consider 

  * `iris` as the small example (7.3 KB)
  * `penguins` from `{palmerpenguins}` as another small-ish example (16.8 KB)
  * `tornados` from `{tidytuesdayR}` 2023-05-16 (2.7 MB)
  
We're going to run these benchmarks on a SLURM cluster, using 16 cores. 

In all of these, we'll use the same library of learners: 

  * `lnr_mean`
  * `lnr_lm`
  * `lnr_rf`
  * `lnr_earth`
  * `lnr_glmnet`
  * `lnr_xgboost`

and their equivalents in the other packages.

# `iris` data

```{r setup}
library(pacman)
p_load('nadir', 'microbenchmark', 'tidytuesdayR', 'future')

# setup multicore use 
future::plan(future::multicore)
petal_formula <- Petal.Width ~ Sepal.Length + Sepal.Width + Petal.Length + Species
```

## `nadir::super_learner()` on `iris` (7.3 KB) data

```{r nadir on iris}
microbenchmark::microbenchmark(
  times = 10,
  list(
    nadir = {
      super_learner(
        data = iris,
        formula = petal_formula,
        learners = list(lnr_mean, lnr_lm, lnr_rf, lnr_earth, lnr_glmnet, lnr_xgboost)
      )
    }
  )
)

microbenchmark::microbenchmark(
  times = 3,
  list(
    nadir_cv = { 
      cv_super_learner(
        data = iris,
        formula = petal_formula,
        learners = list(lnr_mean, lnr_lm, lnr_rf, lnr_earth, lnr_glmnet, lnr_xgboost)
      )
    }
  )
)
```

## `sl3` on `iris` (7.3 KB) data

```{r sl3 on iris}
task <- make_sl3_Task(
  data = iris,
  outcome = "Petal.Width",
  covariates = c("Sepal.Length", 'Sepal.Width', 'Petal.Length', 'Species')
)

lrn_mean <- Lrnr_mean$new()
lrn_lm <- Lrnr_glm$new()
lrn_rf <- Lrnr_randomForest$new()
lrn_earth <- Lrnr_earth$new()
lrn_glmnet <- Lrnr_glmnet$new()
lrn_xgboost <- Lrnr_xgboost$new()

stack <- Stack$new(lrn_mean, lrn_lm, lrn_rf, lrn_earth, lrn_glmnet, lrn_xgboost)

sl <- Lrnr_sl$new(learners = stack, metalearner = Lrnr_nnls$new(),
                  cv_control = list(V = 5))

microbenchmark::microbenchmark(
  times = 10,
  list(
    sl3 = { sl_fit <- sl$train(task = task) }
  )
)

sl_fit <- sl$train(task = task)

system.time({
  sl3_cv = { cv_sl(lrnr_sl = sl_fit, eval_fun = loss_squared_error) }
})
```

## `SuperLearner` on `iris` (7.3 KB) data

```{r SuperLearner on iris}
sl_lib = c( 
  "SL.mean", "SL.lm", "SL.randomForest", "SL.earth", "SL.glmnet", "SL.xgboost")


microbenchmark::microbenchmark(
  times = 10, 
  list(SuperLearner = {
    mcSuperLearner(Y = iris$Petal.Width,
                 X = iris[, -4],
                 SL.library = sl_lib,
                 cvControl = list(V = 5))
  }
  )
)

system.time({
  CV.SuperLearner(
    Y = iris$Petal.Width, 
    X = iris[, -14], 
    SL.library = sl_lib,
    parallel = 'multicore',
    V = 5)
})
```

Cleanup

```{r}
rm(list = ls())
```

# `penguins` data (16.8 KB)

```{r penguins}
penguins <- palmerpenguins::penguins
penguins <- penguins[complete.cases(penguins),]

flipper_length_formula <-
  flipper_length_mm ~ species + island + bill_length_mm +
    bill_depth_mm + body_mass_g + sex
```

## `nadir` on `penguins` data (16.8 KB)

```{r nadir on penguins}
microbenchmark(
  times = 10,
  nadir = {
    nadir::super_learner(
      data = penguins,
      formula = flipper_length_formula,
      learners = list(lnr_mean, lnr_lm, lnr_rf, lnr_earth, lnr_glmnet, lnr_xgboost)
    )
  }
)

microbenchmark::microbenchmark(
  times = 3,
  list(
    nadir_cv = { 
      cv_super_learner(
        data = penguins,
        formula = flipper_length_formula,
        learners = list(lnr_mean, lnr_lm, lnr_rf, lnr_earth, lnr_glmnet, lnr_xgboost)
      )
    }
  )
)
```

## `sl3` on `penguins` data (16.8 KB)

```{r sl3 on penguins}
task <- make_sl3_Task(
  data = penguins,
  outcome = "flipper_length_mm",
  covariates = c("species",
                 "island",
                 "bill_length_mm",
                 "bill_depth_mm",
                 "body_mass_g",
                 "sex")
)

lrn_mean <- Lrnr_mean$new()
lrn_lm <- Lrnr_glm$new()
lrn_rf <- Lrnr_randomForest$new()
lrn_earth <- Lrnr_earth$new()
lrn_glmnet <- Lrnr_glmnet$new()
lrn_xgboost <- Lrnr_xgboost$new()

stack <- Stack$new(lrn_mean, lrn_lm, lrn_rf, lrn_earth, lrn_glmnet, lrn_xgboost)

sl <- Lrnr_sl$new(learners = stack, metalearner = Lrnr_nnls$new(),
                  cv_control = list(V = 5))

microbenchmark::microbenchmark(
  times = 10,
  list(
    sl3 = { sl_fit <- sl$train(task = task) }
  )
)

sl_fit <- sl$train(task = task)

system.time({
  sl3_cv = { cv_sl(lrnr_sl = sl_fit, eval_fun = loss_squared_error) }
})
```

## `SuperLearner` on `penguins` (16.8 KB) data

```{r SuperLearner on penguins}
sl_lib = c( 
  "SL.mean", "SL.lm", "SL.randomForest", "SL.earth", "SL.glmnet", "SL.xgboost")


microbenchmark::microbenchmark(
  times = 10, 
  list(SuperLearner = {
    mcSuperLearner(Y = penguins$flipper_length_mm,
                   X = penguins[, c("species",
                                    "island",
                                    "bill_length_mm",
                                    "bill_depth_mm",
                                    "body_mass_g",
                                    "sex")], 
                 SL.library = sl_lib,
                 cvControl = list(V = 5))
  }
  )
)

num_cores = RhpcBLASctl::get_num_cores()

system.time({
  CV.SuperLearner(
    Y = penguins$flipper_length_mm,
    X = penguins[, c("species",
                     "island",
                     "bill_length_mm",
                     "bill_depth_mm",
                     "body_mass_g",
                     "sex")], 
    SL.library = sl_lib,
    parallel = 'multicore',
    V = 5)
})
```

Cleanup 

```{r}
rm(list=ls())
```


# `tornados` data (2.7 MB)

```{r tornados}
tuesdata <- tidytuesdayR::tt_load('2023-05-16')
tornados <- tuesdata$tornados
tornados <- tornados[,c('yr', 'mo', 'dy', 'mag', 'st', 'inj', 'fat', 'loss')]
tornados <- tornados[complete.cases(tornados),]

# these states appear only very infrequently, like 2 and 1 times respectively — DC and Alaska 
tornados <- tornados |> dplyr::filter(!st %in% c('DC', 'AK'))

tornado_formula <- inj ~ yr + mo + mag + fat + st + loss
```

## `nadir` on `tornados` data (2.7 MB)

```{r nadir on tornados}
system.time({
  super_learner(
    data = tornados,
    formula = tornado_formula,
    learners = list(lnr_mean, lnr_lm, lnr_rf, lnr_earth, lnr_glmnet, lnr_xgboost),
    cv_schema = cv_character_and_factors_schema
  )
})
```

## `sl3` on `tornados` data (2.7 MB)

```{r sl3 on tornados}
task <- make_sl3_Task(
  data = tornados,
  outcome = "inj",
  covariates = c("yr", "mo", "dy", "mag", 
                 "st", "fat", "loss")
)

lrn_mean <- Lrnr_mean$new()
lrn_lm <- Lrnr_glm$new()
lrn_rf <- Lrnr_randomForest$new()
lrn_earth <- Lrnr_earth$new()
lrn_glmnet <- Lrnr_glmnet$new()
lrn_xgboost <- Lrnr_xgboost$new()

stack <- Stack$new(lrn_mean, lrn_lm, lrn_rf, lrn_earth, lrn_glmnet, lrn_xgboost)

sl <- Lrnr_sl$new(learners = stack, metalearner = Lrnr_nnls$new(),
                  cv_control = list(V = 5))

system.time({
  sl_fit <- sl$train(task = task)
})
```

## `SuperLearner` on `tornados` data (2.7 MB)

```{r SuperLearner on tornados}
sl_lib = c( 
  "SL.mean", "SL.lm", "SL.randomForest", "SL.earth", "SL.glmnet", "SL.xgboost")

system.time({
    mcSuperLearner(Y = tornados$inj,
                   X = tornados[, c("yr", "mo", "dy", "mag", 
                 "st", "fat", "loss")], 
                 SL.library = sl_lib,
                 cvControl = list(V = 5))
  })
```
