---
title: "Binary and Multiclass Outcomes"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(nadir)
```

#### I want to use `super_learner()` for binary outcomes.

We'll use the Boston dataset and create a binary outcome for our 
regression problem. 

And of course, we'll train a `super_learner()` to these binary outcomes.  
To handle binary outcomes, we need to adjust the method for 
determining weights. This is because we don't want to use the default `mse()` loss
function, but instead we should to rely on using the negative log likelihood
loss on the held-out data.  To do this appropriately in the context
of binary data, we use the `determine_weights_for_binary_outcomes()` 
function provided by `{nadir}`. 

```{r, class.source='fold-show'}
#| message: false
#| warning: false
#| fig-width: 3
#| fig-height: 3
#| fig-dpi: 300
#| fig-retina: 3

data('Boston', package = 'MASS')

# create a binary outcome to predict
Boston$high_crime <- as.integer(Boston$crim > mean(Boston$crim))
data <- Boston |> dplyr::select(-crim)

# train a super learner on a binary outcome
trained_binary_super_learner <- super_learner(
  data = data,
  formula = high_crime ~ nox + rm + age + tax + ptratio,
  learners = list(
    logistic = lnr_logistic, # the same as a lnr_glm with extra_learner_args 
                             # set to list(family = binomial(link = 'logit'))
                             # for that learner
    rf = lnr_rf_binary,  # random forest
    lm = lnr_lm), # linear probability model
  outcome_type = 'binary',
  verbose = TRUE
)

# let's take a look at the learned weights
trained_binary_super_learner$learner_weights

# what are the predictions? you can think of them as \hat{P}(Y = 1 | X).
# i.e., predictions of P(Y = 1) given X where Y and X are the left & right hand
# side of your regression formula(s)
head(trained_binary_super_learner$sl_predict(data))

# classification table
data.frame(
  truth = data$high_crim, 
  prediction = round(trained_binary_super_learner$sl_predict(data))) |> 
  dplyr::group_by(truth, prediction) |> 
  dplyr::count() |> 
  ggplot2::ggplot(mapping = ggplot2::aes(y = truth, x = prediction, fill = n, label = n)) + 
  ggplot2::geom_tile() + 
  ggplot2::geom_label(fill = 'white', alpha = .7) + 
  ggplot2::scale_fill_distiller(palette = 'Oranges', direction = 1) + 
  ggplot2::scale_x_continuous(breaks = c(0, 1)) + 
  ggplot2::scale_y_continuous(breaks = c(0, 1)) + 
  ggplot2::xlab("super_learner() prediction") + 
  ggplot2::theme_minimal() + 
  ggplot2::theme(legend.position = 'none') + 
  ggplot2::ggtitle("Classification Table")
```

```{r}
#| class-source: 'fold-show'
#| fig-width: 6
#| fig-height: 3
#| fig-dpi: 300
#| fig-retina: 3
data.frame(
  truth = data$high_crim, 
  predicted_pr_of_1 = trained_binary_super_learner$sl_predict(data)) |> 
  ggplot2::ggplot(mapping = ggplot2::aes(x = predicted_pr_of_1)) + 
  ggplot2::geom_histogram() + 
  ggplot2::facet_grid(truth ~ ., labeller = ggplot2::labeller(truth = ~ paste0('truth: ', .))) + 
  ggplot2::theme_bw() + 
  ggplot2::theme(panel.grid.minor = ggplot2::element_blank()) + 
  ggplot2::xlab(bquote(paste("super_learner() predictions, ", hat(bold(P)), '(Y = 1)'))) + 
  ggplot2::ggtitle("Classification Task") 
```

An important thing to know about constructing learners for the binary outcome
context is that `determine_weights_for_binary_outcomes()` requires that the
outputs of the learners on `newdata` are predictions for the outcome being equal
to 1.


## Multiclass Regression, i.e., Multinomial Regression

Earlier we covered binary classification — now we turn to classification 
problems where the dependent variable is one of a discrete number of unique levels. 

`{nadir}` includes two learners that are designed for such multiclass regression
problems: `lnr_multinomial_vglm` and `lnr_multinomial_nnet`. 

We can perform super learning with them and a classification problem like that of 
classifying the penguins' species in the `palmerpenguins` dataset. 

```{r}
#| warning: false
library(palmerpenguins)

df <- penguins[complete.cases(penguins),]

sl_learned_model <- super_learner(
  data = df,
  formulas = list(
    .default = species ~ flipper_length_mm + bill_depth_mm,
    nnet2 = species ~ poly(flipper_length_mm, 2) + poly(bill_depth_mm, 2) + body_mass_g,
    nnet3 = species ~ flipper_length_mm * bill_depth_mm + island
    ),
  learners = list(
    nnet1 = lnr_multinomial_nnet,
    nnet2 = lnr_multinomial_nnet,
    nnet3 = lnr_multinomial_nnet,
    vglm = lnr_multinomial_vglm
    ),
  outcome_type = 'multiclass',
  verbose = TRUE)


compare_learners(sl_learned_model, loss_metric = negative_log_loss)

round(sl_learned_model$learner_weights, 3)
```
