---
title: "Creating Learners"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(nadir)
```

This article contains some advice for writing and constructing new learners. 

### Simple Learner Examples 

```r
lnr_lm <- function(data, formula, ...) {
  model <- stats::lm(formula = formula, data = data, ...)

  predict_from_trained_lm <- function(newdata) {
    predict(model, newdata = newdata, type = 'response')
  }
  return(predict_from_trained_lm)
}
```

### Weights 

We recommend explicitly handling `weights` as an argument to learners so that 
it is a protected argument. Some of the internals of different algorithms may vary,
using other names for `weights` instead, so we recommend doing this to standardize
the weights argument across different learner algorithms.
As a concrete example, the `ranger::ranger()` function takes 
`case.weights` as its argument rather than `weights`.

A typical learner that supports `weights` might look like:
```r
lnr_supportsWeights <- function(data, formula, weights = NULL, ...) {

  # train the model 
  model <- model_fit(data = data, formula = formula, weights = weights, ...)
  
  return(function(newdata) {
    predict(model, newdata = newdata)
  })
}
```


However, some model fitting procedures do not like passing
`weights = NULL` and so it may be necessary to be careful not to pass 
the default `weights = NULL` to the model fitting procedure. 
As an example of this, we refer the curious reader to the source of 
`lnr_glm` in <https://github.com/ctesta01/nadir/blob/main/R/learners.R>.

  model_training_arguments <- list(data = data, formula = formula)
  
  # add weights they aren't missing
  if (! is.null(weights) & length(weights) == nrow(data)) { 
    model_training_arguments$weights <- weights
  }
  
### Attributes

It's recommended that if you create learners, that you also give them a 
couple attributes for a couple of reasons: 

  * If a learner has a `sl_lnr_name` attribute, then this can be automatically used in the outputs if a name for the learner is left unspecified. 
  * If a learner has a `sl_lnr_type` attribute, it will be checked against the `output_type` argument to `super_learner()`.

To set these attributes, when making a new learner, one should run something 
along the lines of 
```{r}
#| eval: false
lnr_myNewLearner <- function(data, formula, ...) {
  model <- # fit your learner given data, formula, ...
    
  predictor_fn <- function(newdata) {
    predict(model, newdata = newdata)
  }
  return(predictor)
}
attr(lnr_myNewLearner, 'sl_lnr_name') <- 'newLearnerName'
attr(lnr_myNewLearner, 'sl_lnr_type') <- 'continuous' # or c('continuous', 'binary') and similar
# see ?nadir_supported_types
```
