---
title: "Creating Learners"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(nadir)
```

This article contains some advice for writing and constructing new learners. 

### Simple Learner Examples 

```r
lnr_lm <- function(data, formula, ...) {
  model <- stats::lm(formula = formula, data = data, ...)

  predict_from_trained_lm <- function(newdata) {
    predict(model, newdata = newdata, type = 'response')
  }
  return(predict_from_trained_lm)
}
```

<!-- 
### Weights 

We recommend explicitly handling `weights` as an argument to learners so that 
it is a protected argument. Some of the internals of different algorithms may vary,
using other names for `weights` instead, so we recommend doing this to standardize
the weights argument across different learner algorithms.

Still trying to figure out how to handle weights.
-->

### Attributes

It's recommended that if you create learners, that you also give them a 
couple attributes for a couple of reasons: 

  * If a learner has a `sl_lnr_name` attribute, then this can be automatically used in the outputs if a name for the learner is left unspecified. 
  * If a learner has a `sl_lnr_type` attribute, it will be checked against the `output_type` argument to `super_learner()`.

To set these attributes, when making a new learner, one should run something 
along the lines of 
```{r}
#| eval: false
lnr_myNewLearner <- function(data, formula, weights, ...) {
  model <- # fit your learner given data, formula, weights, ...
}
```
