---
title: "Cross-Fitting and Out-of-Fold Prediction"
editor_options: 
  chunk_output_type: inline
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE}
library(nadir)
set.seed(60615)
```

## Why cross-fit?

`super_learner()` already uses cross-validation internally: each candidate
learner is trained on each fold's training split and scored on its held-out
split, and the ensemble weights are chosen to minimize held-out loss. So the
*learners* never score their own training data. But the ensemble *weights*
are estimated from the pooled held-out predictions of every observation —
including observation $i$'s. For most predictive uses this is fine. For
semiparametric causal inference (AIPW, TMLE, double machine learning), the
theory asks for a little more: each observation's nuisance prediction
$\hat m(X_i)$ should come from a model fit entirely without observation $i$ —
weights included.

`crossfit_super_learner()` provides exactly that. It splits the data into
`n_folds` outer folds and fits a *complete* `super_learner()` — candidate
learners and ensemble weights, via its own `inner_n_folds`-fold
cross-validation — on each outer training split, then predicts on the
corresponding held-out fold. Every out-of-fold prediction therefore comes
from an ensemble whose base learners *and* meta-learned weights never saw
that row.

The three fitting functions answer three different questions:

- `super_learner()`: "give me the best predictor of $Y$" — use `$predict()`
  on new data;
- `crossfit_super_learner()`: "give me honest per-observation predictions
  for my training rows" — there deliberately is no single `$predict()`
  (calling it errors with directions), because a cross-fitted model *is*
  its collection of fold-specific fits;
- `cv_super_learner()`: "how good is the super learning procedure itself?"
  — an honest estimate of the whole procedure's held-out loss (it is built
  on the crossfit machinery; see `$crossfit`).

## A first cross-fitted model

```{r first-crossfit}
cf <- crossfit_super_learner(
  data = mtcars,
  formulas = mpg ~ hp + wt + am,
  learners = list(lm = lnr_lm, mean = lnr_mean),
  n_folds = 5
)
cf
```

The honest predictions live in `$oof_predictions`, a plain numeric vector in
the row order of the input data:

```{r oof-vector}
head(cf$oof_predictions)
cf$cv_loss # empirical loss of the cross-fitted predictions
```

Because each outer fold estimates its own ensemble, `coef()` returns a
folds-by-learners weight matrix, and `summary()` reports how stable those
weights are across folds — a useful diagnostic before trusting the
cross-fitted predictions downstream:

```{r coefs-summary}
coef(cf)
summary(cf)
```

`fitted(cf)` and `residuals(cf)` are the familiar S3 spellings of the same
quantities: `fitted()` is `$oof_predictions`, and `residuals()` is observed
minus out-of-fold predicted, both in input row order.

## The `oof_predict` family

The same four-method interface is available on both `super_learner()` fits
(class `nadir_sl_model`) and `crossfit_super_learner()` fits (class
`nadir_crossfit_sl`):

- **`$oof_predictions`** — the stored vector of out-of-fold predictions, in
  input row order. Rows never held out by a non-covering `cv_schema` are
  `NA`.
- **`$oof_predict(newdata, rowids = NULL)`** — out-of-fold predictions for
  (possibly modified) versions of the *training* rows: row $i$ of `newdata`
  is predicted by the fold-specific fit that never saw row $i$. With no
  arguments it returns the stored vector.
- **`$oof_predict_modified(modify)`** — the counterfactual workhorse:
  applies `modify(newdata)` to each held-out fold before predicting.
  `$oof_predict_modified(NULL)` re-predicts the unmodified data and agrees
  with `$oof_predictions`.
- **`$oof_predict_fold(newdata_list)`** — per-fold prediction vectors, for
  workflows that need the fold structure explicitly.

A note on row matching in `$oof_predict()`: out-of-fold prediction is only
meaningful for rows the model was cross-fit on, so `newdata` rows must be
matched back to training rows. If you supplied `rowids` at fit time,
matching is by id — shuffled or subset `newdata` is fine, and `rowids` is
then required at predict time. Without fit-time `rowids`, rows are matched
by position and a one-time warning reminds you of the assumption. We highly 
recommend using the rowids.

```{r rowids}
cf_ids <- crossfit_super_learner(
  data = mtcars,
  formulas = mpg ~ hp + wt + am,
  learners = list(lm = lnr_lm, mean = lnr_mean),
  n_folds = 5,
  rowids = rownames(mtcars)
)

# a shuffled subset, matched by id for demonstration purposes
shuffled <- mtcars[c("Valiant", "Fiat 128", "Lotus Europa"), ]
cf_ids$oof_predict(shuffled, rowids = rownames(shuffled))
```

## Honest nuisances for causal inference

The point of all this machinery is estimators built from cross-fitted
nuisance functions. Here is a complete AIPW estimate of an average treatment
effect: an outcome model $\hat m(A, W)$ and a propensity model
$\hat g(W)$, each cross-fit, with the counterfactual predictions
$\hat m(1, W_i)$ and $\hat m(0, W_i)$ obtained through
`$oof_predict_modified()`.

```{r aipw}
n <- 800
W1 <- rnorm(n)
W2 <- rnorm(n)
A <- rbinom(n, 1, plogis(0.4 * W1 - 0.3 * W2))
Y <- 1 + A + 0.8 * W1 - 0.5 * W2 + rnorm(n) # true ATE = 1
d <- data.frame(W1 = W1, W2 = W2, A = A, Y = Y)

cf_Q <- crossfit_super_learner(
  data = d, formulas = Y ~ A + W1 + W2,
  learners = list(lm = lnr_lm, mean = lnr_mean),
  n_folds = 5
)
cf_g <- crossfit_super_learner(
  data = d, formulas = A ~ W1 + W2,
  learners = list(logistic = lnr_logistic),
  outcome_type = "binary",
  n_folds = 5
)

bound <- function(x, l = 0.005) pmin(pmax(x, l), 1 - l)

QAW <- cf_Q$oof_predictions # m(A_i, W_i)
Q1W <- cf_Q$oof_predict_modified(function(dd) {
  dd$A <- 1
  dd
}) # m(1, W_i)
Q0W <- cf_Q$oof_predict_modified(function(dd) {
  dd$A <- 0
  dd
}) # m(0, W_i)
gW <- bound(cf_g$oof_predictions) # g(W_i)

H <- A / gW - (1 - A) / (1 - gW)
phi <- (Q1W - Q0W) + H * (Y - QAW) # efficient influence function + psi

psi <- mean(phi)
se <- sd(phi) / sqrt(n)
c(estimate = psi, lower = psi - 1.96 * se, upper = psi + 1.96 * se)
```

Every quantity entering `phi` for observation $i$ was produced by fits —
learners and ensemble weights both — which were estimated without observation $i$,
and that is what makes the nice efficient-influence-function based standard error
valid to use for inference.

Two refinements are provided for production use. 
First, estimators like CV-TMLE want the
outcome and propensity models cross-fit on the *same* folds; pass both fits
a shared custom `cv_schema` that fixes one fold assignment (the package's
test suite contains a complete shared-fold CV-TMLE recipe). Second, if you
are fitting `super_learner()` purely for its out-of-fold interfaces,
`train_on_whole_dataset = FALSE` skips the final whole-data learner fits —
saving roughly `1/n_folds` of the compute — while keeping
`$oof_predictions`, `$oof_predict()`, `$oof_predict_modified()`, and
`$oof_predict_fold()` fully functional and available for use.
