---
title: "Density Estimation"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(nadir)
library(ggplot2)
```

<!-- 
NOTES FOR AUTHORS:
Given the intent to apply super_learner() in causal settings, would it be
useful to go through this replacing "outcome" with "exposure/treatment" 
and add a notice at the top? 
Similarly, predictors == covariates or confounders. 
-->

Thinking about applications of the super learner algorithm 
to causal inference, one key application is through modeling the treatment/exposure 
mechanism. 

In many cases, the treatment/exposure is continuous, necessitating the 
estimation of a "generalized" propensity score — generalized in the sense that 
the treatment/exposure is no longer binary but continuous, and hence
our propensity score model is a probability density model for a continuous range
of exposure values rather than just the probability 
of a binary treatment/no-treatment variable. 

Here we demonstrate the application of `nadir::super_learner()` in estimating
such generalized propensity scores. 

The following examples are provided in the spirit of showing a myriad number
of ways `{nadir}` can be used to tackle density estimation. 

## Basic Syntax for Fitting Density Models

To fit a conditional density model, we need a few things: 

  * a dataset
  * to select the outcome, and predictor variables (which together define a formula)
  * a collection of candidate learners
  * and to use the negative log loss function in the super learner algorithm for
  choosing among the learners.

```{r}
# specify our data and regression problem 
# ---------------------------------------

# in order to build a weighting based estimator, we might fit a conditional
# density model of our continuous exposure 
data("Boston", package = "MASS")

# we're going to regress the crime measurement variable in the Boston dataset
# on all the other variables. 
# 
# and we want to control for the following potential confounders (all other variables in the Boston dataset) 
# 
# so we want to regress crime on all the other variables in the Boston dataset 
reg_formula <- crim ~ . 


# specify a collection of candidate learners
# ------------------------------------------

# we'll use a normal error model with a linear conditional mean model, 
# and random forest, ranger, lm, and earth with kernel bandwidth density() models for the 
# error distribution

# nadir provides the helper function lnr_homoskedastic_density that can be used
# to build a homoskedastic density learner out of a given conditional mean learner.

lnr_rf_homoskedastic_density <- function(data, formula, ...) { 
  lnr_homoskedastic_density(data, formula, mean_lnr = lnr_rf, ...)
}

lnr_ranger_homoskedastic_density <- function(data, formula, ...) { 
  lnr_homoskedastic_density(data, formula, mean_lnr = lnr_ranger, ...)
}

lnr_lm_homoskedastic_density <- function(data, formula, ...) { 
  lnr_homoskedastic_density(data, formula, mean_lnr = lnr_lm, ...)
}

lnr_earth_homoskedastic_density <- function(data, formula, ...) { 
  lnr_homoskedastic_density(data, formula, mean_lnr = lnr_earth, ...)
}

# we recommend setting the sl_lnr_type attribute to avoid warnings from super_learner()
# 
# this tells the super_learner() function that these learners are designed
# to work with density estimation problems.
# 
# without this, users will see a friendly warning asking them if they're sure 
# they want to use the given learners with the outcome_type = 'density'.
# 
attr(lnr_rf_homoskedastic_density, 'sl_lnr_type') <- 'density'
attr(lnr_ranger_homoskedastic_density, 'sl_lnr_type') <- 'density'
attr(lnr_lm_homoskedastic_density, 'sl_lnr_type') <- 'density'
attr(lnr_earth_homoskedastic_density, 'sl_lnr_type') <- 'density'


# super learning: 
# choosing weights for candidate learners and ensembling
# ======================================================

learned_sl_density_model <- super_learner(
  data = Boston,
  formula = reg_formula,
  learners = list(
    normal = lnr_lm_density,
    ranger = lnr_ranger_homoskedastic_density,
    earth = lnr_earth_homoskedastic_density,
    lm = lnr_lm_homoskedastic_density
    ),
  outcome_type = 'density',
  verbose = TRUE
)

# if you prefer not to code up the new learners as above, another way to perform
# the same super learning as above is to use the `extra_learner_args` option
# with named learners as follows:
#
# learned_sl_density_model <- super_learner(
#   data = Boston,
#   formula = reg_formula,
#   learners = list(
#     normal = lnr_lm,
#     ranger = lnr_homoskedastic_density,
#     earth = lnr_homoskedastic_density,
#     lm = lnr_lm_homoskedastic_density
#     ),
#   extra_learner_args = list(
#     ranger = list(mean_lnr = lnr_ranger),
#     earth = list(mean_lnr = lnr_earth),
#     lm = list(mean_lnr = lnr_lm)),
#   outcome_type = 'density',
#   verbose = TRUE
# )
# 
# one reason it can be useful to create the learners explicitly is to be able to
# visualize and debug their behaviors individually.


# fit diagnostics
# ===============

# two things we might want to do with a fit super learner model are:
# 1) see how each candidate learner performed with regard to the specified loss
#    function
# 2) see the weights assigned to each learner (how favored they are in the final
#    ensemble model).
#
# to be able to see these diagnostics, it's important that super_learner() is
# run with vebose = TRUE otherwise these extra details are not stored in the 
# output object. 

# 1) compare the learners using negative log likelihood loss
compare_learners(learned_sl_density_model, loss_metric = negative_log_loss)

# 2) see the weights given to each candidate learner
learned_sl_density_model$learner_weights
```


<details>
<summary>
Validating that conditional density estimation is working correctly
</summary>
## Let's validate that conditional density works the way it should 

One test we can do to check that conditional density estimation 
is working correctly is that once we condition on a set of predictors, 
the produced probability density function should integrate to 1 -- 
that is, its area under the curve should be equal to 1. 

We show an example of a test here: 

```{r}
lm_density_predict <- lnr_lm_homoskedastic_density(Boston, reg_formula)

f_lm <- function(ys) {
  x <- Boston[1,]
  sapply(ys, function(y) {
    x[['crim']] <- y
    lm_density_predict(x)
  })
}

integrate(f_lm, min(Boston$crim) - sd(Boston$crim), max(Boston$crim) + sd(Boston$crim), subdivisions = 10000)


earth_density_predict <- lnr_earth_homoskedastic_density(Boston, reg_formula, density_args = list(bw = 30))

f_earth <- function(ys) {
  x <- Boston[1,]
  sapply(ys, function(y) {
    x[['crim']] <- y
    earth_density_predict(x)
  })
}
earth_density_predict(Boston[1,])

# we're looking for this number to be very close to 1:
integrate(f_earth, min(Boston$crim) - 10*sd(Boston$crim), max(Boston$crim) + 10*sd(Boston$crim))


# sometimes if integrate() has trouble, it's useful to perform the integration by hand:
y_seq <- seq(min(Boston$crim) - 10*sd(Boston$crim), max(Boston$crim) + 10*sd(Boston$crim), length.out = 10000)
delta_y <- y_seq[2]-y_seq[1]
sum(f_earth(y_seq)*delta_y) # we're looking for this to be close to 1
```
</details>

## Heteroskedastic Learners 

```{r}
#| warning: false
#| message: false
lnr_earth_mean_glm_var_heteroskedastic_density <- function(data, formula, ...) { 
  lnr_heteroskedastic_density(
    data,
    formula,
    mean_lnr = lnr_earth,
    var_lnr = lnr_lm,
    density_args = list(bw=5),
    ...
  )
}

Boston$log_crim <- log(Boston$crim)

earth_mean_glm_var_heteroskedastic_predict <-
  lnr_earth_mean_glm_var_heteroskedastic_density(Boston[,-1], log_crim ~ medv)

earth_mean_glm_var_heteroskedastic_predict(Boston[1,])

f_earth_glm <- function(ys) {
  x <- Boston[2,]
  sapply(ys, function(y) {
    x[['log_crim']] <- y
    earth_mean_glm_var_heteroskedastic_predict(x)
  })
}

# check that the estimated density has area-under-the-curve = 1
integrate(
  f_earth_glm,
  min(Boston$log_crim) - 10 * sd(Boston$log_crim),
  max(Boston$log_crim) + 10 * sd(Boston$log_crim),
  subdivisions = 1000
)

# create a grid to predict densities over
prediction_grid <- expand.grid(
  log_crim = log(seq(min(Boston$crim), max(Boston$crim), length.out = 100)),
  medv = seq(min(Boston$medv), max(Boston$medv), length.out = 100))

# calculate densities for every grid-point
prediction_grid$density_prediction <- earth_mean_glm_var_heteroskedastic_predict(
  prediction_grid)

# visualize the density model across a grid of outcomes and predictors
ggplot2::ggplot(Boston, aes(x = medv, y = exp(log_crim))) + 
  geom_tile(
    data = prediction_grid, 
    mapping = aes(fill = density_prediction)) + 
  geom_point() + 
  theme_minimal() + 
  MetBrewer::scale_fill_met_c('Hiroshige', direction = -1) + 
  ggtitle("Heteroskedastic Density Estimation Model", 
          "(Mean model: earth, Variance model: GLM)")
```

The above example is in 2D (e.g., one predictor and one outcome variable), 
but in practical settings we may be estimating the density for one variable
conditional on many other variables.  The syntax easily extends as follows,
recalling that `reg_formula` was instantiated above as `r reg_formula`.

```{r}
#| message: false
#| warning: false
lnr_earth_glm_heteroskedastic_density <- function(data, formula, ...) { 
  lnr_heteroskedastic_density(data, formula, mean_lnr = lnr_earth, 
                            var_lnr = lnr_mean, 
                            density_args = list(bw = 3),
                            ...)
}

lnr_earth_glm_heteroskedastic_predict <-
  lnr_earth_glm_heteroskedastic_density(Boston, reg_formula)

lnr_earth_glm_heteroskedastic_predict(Boston[1,])

f_earth_mean <- function(ys) {
  x <- Boston[1,]
  sapply(ys, function(y) {
    x[['crim']] <- y
    lnr_earth_glm_heteroskedastic_predict(x)
  })
}

integrate(f_earth, min(Boston$crim) - sd(Boston$crim), max(Boston$crim) + sd(Boston$crim))
```

We've omitted visualizing what the predictions look like for this model. In general, 
while it can be useful to spot check whether or not integration is working correctly,
or if certain heuristics hold, it can be hard to reason about how a 2D slice of a higher
dimensional conditional density model should appear. 

For example, some intuition that begins to fail in higher dimensional conditional
density models is that the density should be higher where the data points are clustered: 
if one conditions on a set of all but 2 parameter values and visualizes a grid like 
in the above example for the remaining predictor and outcome variables, the 
set of variables conditioned on may be more consistent with higher density in a 
region outside the bulk of the data distribution if the conditioning set is more
consistent with outliers. 

In general, this is why for higher dimensional models, we prefer to rely on 
checks like "does the density integrate to 1?" after conditioning on a full 
set of predictors and integrating over the possible values of the outcome.

## 2D Density Estimation 

Perhaps visualizing density estimation in 2-to-3 dimensions will help 
people understand what's going on. 

Suppose our regression problem is to regress flipper length on bill depth from
the palmer penguins dataset.  We'll fit some conditional density models and 
visualize their predictions. 

```{r}
#| fig-width: 8
#| fig-height: 5
#| fig-dpi: 300
#| fig-retina: 3

library(palmerpenguins)
library(MetBrewer)

data <- penguins[,c('flipper_length_mm', 'bill_depth_mm')]
data <- data[complete.cases(data),]

# conditional normal example:

learned_conditional_normal <- lnr_lm_density(data, flipper_length_mm ~ bill_depth_mm)

prediction_grid <- expand.grid(
  flipper_length_mm = seq(172, 231, length.out = 100),
  bill_depth_mm = seq(13.1, 21.5, length.out = 100))

prediction_grid$density_prediction <- learned_conditional_normal(prediction_grid)

ggplot(data, aes(x = bill_depth_mm, y = flipper_length_mm)) + 
  geom_tile(
    data = prediction_grid, 
    mapping = aes(fill = density_prediction)) + 
  geom_point() + 
  theme_minimal() + 
  MetBrewer::scale_fill_met_c('Hiroshige', direction = -1) + 
  ggtitle("Conditionally Normal Density Model")

# earth and stats::density

learned_earth_homoskedastic_density <-
  lnr_earth_homoskedastic_density(data, flipper_length_mm ~ bill_depth_mm)

prediction_grid$density_prediction <- learned_earth_homoskedastic_density(prediction_grid)

ggplot(data, aes(x = bill_depth_mm, y = flipper_length_mm)) + 
  geom_tile(
    data = prediction_grid, 
    mapping = aes(fill = density_prediction)) + 
  geom_point() + 
  theme_minimal() + 
  MetBrewer::scale_fill_met_c('Hiroshige', direction = -1) + 
  ggtitle("Using earth::earth for conditional mean and stats::density for error")

# randomForest and stats::density

learned_rf_homoskedastic_density <-
  lnr_rf_homoskedastic_density(data, flipper_length_mm ~ bill_depth_mm)

prediction_grid$density_prediction <- learned_rf_homoskedastic_density(prediction_grid)

ggplot(data, aes(x = bill_depth_mm, y = flipper_length_mm)) + 
  geom_tile(
    data = prediction_grid, 
    mapping = aes(fill = density_prediction)) + 
  geom_point() + 
  theme_minimal() + 
  MetBrewer::scale_fill_met_c('Hiroshige', direction = -1) + 
  ggtitle("Using randomForest::randomForest for conditional mean and stats::density for error")

# ranger and stats::density

learned_ranger_homoskedastic_density <-
  lnr_ranger_homoskedastic_density(data, flipper_length_mm ~ bill_depth_mm)

prediction_grid$density_prediction <- learned_ranger_homoskedastic_density(prediction_grid)

ggplot(data, aes(x = bill_depth_mm, y = flipper_length_mm)) + 
  geom_tile(
    data = prediction_grid, 
    mapping = aes(fill = density_prediction)) + 
  geom_point() +
  theme_minimal() + 
  MetBrewer::scale_fill_met_c('Hiroshige', direction = -1) + 
  ggtitle("Using ranger::ranger for conditional mean and stats::density for error")

# glm and stats::density

learned_glm_homoskedastic_density <-
  lnr_homoskedastic_density(data, flipper_length_mm ~ bill_depth_mm, mean_lnr = lnr_glm)

prediction_grid$density_prediction <- learned_glm_homoskedastic_density(prediction_grid)

ggplot(data, aes(x = bill_depth_mm, y = flipper_length_mm)) + 
  geom_tile(
    data = prediction_grid, 
    mapping = aes(fill = density_prediction)) + 
  geom_point() + 
  theme_minimal() + 
  MetBrewer::scale_fill_met_c('Hiroshige', direction = -1) + 
  ggtitle("Using stats::glm for conditional mean and stats::density for error")


#  heteroskedastic method 

learned_earth_heteroskedastic_density <-
  lnr_heteroskedastic_density(data,
                              flipper_length_mm ~ bill_depth_mm,
                              mean_lnr = lnr_earth,
                              var_lnr = lnr_glm)

prediction_grid$density_prediction <- learned_earth_heteroskedastic_density(prediction_grid)

ggplot(data, aes(x = bill_depth_mm, y = flipper_length_mm)) + 
  geom_tile(
    data = prediction_grid, 
    mapping = aes(fill = density_prediction)) + 
  geom_point() + 
  theme_minimal() + 
  MetBrewer::scale_fill_met_c('Hiroshige', direction = -1) + 
  ggtitle("Using earth::earth for conditional mean, stats::density for error distribution, \nand a stats::glm for predicting heteroskedasticity")

# super learner

learned_sl_density <-
  super_learner(data,
    flipper_length_mm ~ bill_depth_mm,
    learners = list(
      lm = lnr_lm_density,
      earth = lnr_earth_homoskedastic_density,
      rf = lnr_rf_homoskedastic_density,
      ranger = lnr_ranger_homoskedastic_density,
      heteroskedastic = lnr_heteroskedastic_density
      ),
    extra_learner_args = list(
      heteroskedastic = list(mean_lnr = lnr_earth, var_lnr = lnr_glm)
      ),
    determine_super_learner_weights = determine_weights_using_neg_log_loss)

prediction_grid$density_prediction <- learned_sl_density(prediction_grid)

ggplot(data, aes(x = bill_depth_mm, y = flipper_length_mm)) + 
  geom_tile(
    data = prediction_grid, 
    mapping = aes(fill = density_prediction)) + 
  geom_point() + 
  theme_minimal() + 
  MetBrewer::scale_fill_met_c('Hiroshige', direction = -1) + 
  ggtitle("Using super_learner on all of the above with negative log loss")
```
