---
title: "Density Estimation"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(nadir)
```

Thinking about applications of the super learner algorithm 
to causal inference, one key application is through modeling the treatment/exposure 
mechanism. 

In many cases, the treatment/exposure is continuous, necessitating the 
estimation of a "generalized" propensity score — generalized in the sense that 
the treatment/exposure is no longer binary but continuous. 

Here we demonstrate the application of `nadir::super_learner()` in estimating
such a generalized propensity score.

We're not going to worry (yet) about having the most realistic examples; the 
point is that we want to see if we can get density estimation working. 

```{r}
# in order to build a weighting based estimator, we might fit a conditional
# density model of our continuous exposure 
data("Boston", package = "MASS")

# suppose our hypothesis of interest is "does crime reduce median home value?" 
# and we want to control for the following potential confounders (all other variables in the Boston dataset) 
# 
# so we want to regress crime on all the other variables in the Boston dataset 
reg_formula <- as.formula(paste0("crim ~ ", paste0(setdiff(colnames(Boston), c('medv', 'crim')), collapse = ' + ')))

# we would want to regress poverty on these variables

lnr_rf_homoskedastic_density <- function(data, formula, ...) { 
  lnr_homoskedastic_density(data, formula, mean_lnr = lnr_rf, ...)
}

lnr_ranger_homoskedastic_density <- function(data, formula, ...) { 
  lnr_homoskedastic_density(data, formula, mean_lnr = lnr_ranger, ...)
}

lnr_lm_homoskedastic_density <- function(data, formula, ...) { 
  lnr_homoskedastic_density(data, formula, mean_lnr = lnr_lm, ...)
}

lnr_earth_homoskedastic_density <- function(data, formula, ...) { 
  lnr_homoskedastic_density(data, formula, mean_lnr = lnr_earth, ...)
}
# we recommend setting the sl_lnr_type attribute to avoid warnings from super_learner()
attr(lnr_rf_homoskedastic_density, 'sl_lnr_type') <- 'density'
attr(lnr_ranger_homoskedastic_density, 'sl_lnr_type') <- 'density'
attr(lnr_lm_homoskedastic_density, 'sl_lnr_type') <- 'density'
attr(lnr_earth_homoskedastic_density, 'sl_lnr_type') <- 'density'

learned_sl_density_model <- super_learner(
  data = Boston,
  formula = reg_formula,
  learners = list(
    normal = lnr_lm_density,
    ranger = lnr_ranger_homoskedastic_density,
    earth = lnr_earth_homoskedastic_density,
    lm = lnr_lm_homoskedastic_density
    ),
  outcome_type = 'density',
  verbose = TRUE
)

# compare the learners using negative log likelihood loss
compare_learners(learned_sl_density_model, loss_metric = negative_log_loss)

learned_sl_density_model$learner_weights
```



## Let's validate that conditional density works the way it should 

```{r}


lm_density_predict <- lnr_lm_homoskedastic_density(Boston, reg_formula)

f_lm <- function(ys) {
  x <- Boston[1,]
  sapply(ys, function(y) {
    x[['crim']] <- y
    lm_density_predict(x)
  })
}

integrate(f_lm, min(Boston$crim) - sd(Boston$crim), max(Boston$crim) + sd(Boston$crim), subdivisions = 10000)


earth_density_predict <- lnr_earth_homoskedastic_density(Boston, reg_formula, density_args = list(bw = 30))

f_earth <- function(ys) {
  x <- Boston[1,]
  sapply(ys, function(y) {
    x[['crim']] <- y
    earth_density_predict(x)
  })
}
earth_density_predict(Boston[1,])

integrate(f_earth, min(Boston$crim) - 10*sd(Boston$crim), max(Boston$crim) + 10*sd(Boston$crim))

y_seq <- seq(min(Boston$crim) - 10*sd(Boston$crim), max(Boston$crim) + 10*sd(Boston$crim), length.out = 10000)
# f_earth(y_seq)

delta_y <- y_seq[2]-y_seq[1]

sum(f_earth(y_seq)*delta_y)

```

## Heteroskedastic Learners 

```{r}
#| warning: false
#| message: false
lnr_earth_mean_glm_var_heteroskedastic_density <- function(data, formula, ...) { 
  lnr_heteroskedastic_density(data, formula, mean_lnr = lnr_earth, 
                            var_lnr = lnr_glm, 
                            var_lnr_args = list(family = gaussian(link = 'log')),
                            ...)
}

earth_mean_glm_var_heteroskedastic_predict <- lnr_earth_mean_glm_var_heteroskedastic_density(Boston, reg_formula)

earth_mean_glm_var_heteroskedastic_predict(Boston[1,])

f_earth_glm <- function(ys) {
  x <- Boston[1,]
  sapply(ys, function(y) {
    x[['crim']] <- y
    earth_mean_glm_var_heteroskedastic_predict(x)
  })
}

integrate(f_earth_glm, min(Boston$crim) - 10*sd(Boston$crim), max(Boston$crim) + 10*sd(Boston$crim), subdivisions = 1000)

sum(f_earth_glm(y_seq)*delta_y)

```

```{r}
#| message: false
#| warning: false
lnr_earth_glm_heteroskedastic_density <- function(data, formula, ...) { 
  lnr_heteroskedastic_density(data, formula, mean_lnr = lnr_earth, 
                            var_lnr = lnr_mean, 
                            density_args = list(bw = 3),
                            ...)
}

lnr_earth_glm_heteroskedastic_predict <- lnr_earth_glm_heteroskedastic_density(Boston, reg_formula)

lnr_earth_glm_heteroskedastic_predict(Boston[1,])

f_earth_mean <- function(ys) {
  x <- Boston[1,]
  sapply(ys, function(y) {
    x[['crim']] <- y
    lnr_earth_glm_heteroskedastic_predict(x)
  })
}

# integrate(f_earth, min(Boston$crim) - sd(Boston$crim), max(Boston$crim) + sd(Boston$crim), subdivisions = )

sum(f_earth_mean(y_seq)*delta_y)
```

This indicates that as of now, the `lnr_heteroskedastic_density()` method is not working
correctly.




## 2D Density Estimation 

Perhaps visualizing density estimation in 2-to-3 dimensions will help 
people understand what's going on. 

Suppose our regression problem is to regress flipper length on bill depth from
the palmer penguins dataset.  We'll fit some conditional density models and 
visualize their predictions. 

```{r}
#| fig-width: 8
#| fig-height: 5
#| fig-dpi: 300
#| fig-retina: 3

library(palmerpenguins)
library(ggplot2)
library(MetBrewer)

data <- penguins[,c('flipper_length_mm', 'bill_depth_mm')]
data <- data[complete.cases(data),]

# conditional normal example:

learned_conditional_normal <- lnr_lm_density(data, flipper_length_mm ~ bill_depth_mm)

prediction_grid <- expand.grid(
  flipper_length_mm = seq(172, 231, length.out = 100),
  bill_depth_mm = seq(13.1, 21.5, length.out = 100))

prediction_grid$density_prediction <- learned_conditional_normal(prediction_grid)

ggplot(data, aes(x = bill_depth_mm, y = flipper_length_mm)) + 
  geom_tile(
    data = prediction_grid, 
    mapping = aes(fill = density_prediction)) + 
  geom_point() + 
  theme_minimal() + 
  MetBrewer::scale_fill_met_c('Hiroshige', direction = -1) + 
  ggtitle("Conditionally Normal Density Model")

# earth and stats::density

learned_earth_homoskedastic_density <-
  lnr_earth_homoskedastic_density(data, flipper_length_mm ~ bill_depth_mm)

prediction_grid$density_prediction <- learned_earth_homoskedastic_density(prediction_grid)

ggplot(data, aes(x = bill_depth_mm, y = flipper_length_mm)) + 
  geom_tile(
    data = prediction_grid, 
    mapping = aes(fill = density_prediction)) + 
  geom_point() + 
  theme_minimal() + 
  MetBrewer::scale_fill_met_c('Hiroshige', direction = -1) + 
  ggtitle("Using earth::earth for conditional mean and stats::density for error")

# randomForest and stats::density

learned_rf_homoskedastic_density <-
  lnr_rf_homoskedastic_density(data, flipper_length_mm ~ bill_depth_mm)

prediction_grid$density_prediction <- learned_rf_homoskedastic_density(prediction_grid)

ggplot(data, aes(x = bill_depth_mm, y = flipper_length_mm)) + 
  geom_tile(
    data = prediction_grid, 
    mapping = aes(fill = density_prediction)) + 
  geom_point() + 
  theme_minimal() + 
  MetBrewer::scale_fill_met_c('Hiroshige', direction = -1) + 
  ggtitle("Using randomForest::randomForest for conditional mean and stats::density for error")

# ranger and stats::density

learned_ranger_homoskedastic_density <-
  lnr_ranger_homoskedastic_density(data, flipper_length_mm ~ bill_depth_mm)

prediction_grid$density_prediction <- learned_ranger_homoskedastic_density(prediction_grid)

ggplot(data, aes(x = bill_depth_mm, y = flipper_length_mm)) + 
  geom_tile(
    data = prediction_grid, 
    mapping = aes(fill = density_prediction)) + 
  geom_point() +
  theme_minimal() + 
  MetBrewer::scale_fill_met_c('Hiroshige', direction = -1) + 
  ggtitle("Using ranger::ranger for conditional mean and stats::density for error")

# glm and stats::density

learned_glm_homoskedastic_density <-
  lnr_homoskedastic_density(data, flipper_length_mm ~ bill_depth_mm, mean_lnr = lnr_glm)

prediction_grid$density_prediction <- learned_glm_homoskedastic_density(prediction_grid)

ggplot(data, aes(x = bill_depth_mm, y = flipper_length_mm)) + 
  geom_tile(
    data = prediction_grid, 
    mapping = aes(fill = density_prediction)) + 
  geom_point() + 
  theme_minimal() + 
  MetBrewer::scale_fill_met_c('Hiroshige', direction = -1) + 
  ggtitle("Using stats::glm for conditional mean and stats::density for error")


#  heteroskedastic method 

learned_earth_heteroskedastic_density <-
  lnr_heteroskedastic_density(data,
                              flipper_length_mm ~ bill_depth_mm,
                              mean_lnr = lnr_earth,
                              var_lnr = lnr_glm)

prediction_grid$density_prediction <- learned_earth_heteroskedastic_density(prediction_grid)

ggplot(data, aes(x = bill_depth_mm, y = flipper_length_mm)) + 
  geom_tile(
    data = prediction_grid, 
    mapping = aes(fill = density_prediction)) + 
  geom_point() + 
  theme_minimal() + 
  MetBrewer::scale_fill_met_c('Hiroshige', direction = -1) + 
  ggtitle("Using earth::earth for conditional mean, stats::density for error distribution, \nand a stats::glm for predicting heteroskedasticity")

# super learner

learned_sl_density <-
  super_learner(data,
    flipper_length_mm ~ bill_depth_mm,
    learners = list(
      lm = lnr_lm_density,
      earth = lnr_earth_homoskedastic_density,
      rf = lnr_rf_homoskedastic_density,
      ranger = lnr_ranger_homoskedastic_density,
      heteroskedastic = lnr_heteroskedastic_density
      ),
    extra_learner_args = list(
      heteroskedastic = list(mean_lnr = lnr_earth, var_lnr = lnr_glm)
      ),
    determine_super_learner_weights = determine_weights_using_neg_log_loss)

prediction_grid$density_prediction <- learned_sl_density(prediction_grid)

ggplot(data, aes(x = bill_depth_mm, y = flipper_length_mm)) + 
  geom_tile(
    data = prediction_grid, 
    mapping = aes(fill = density_prediction)) + 
  geom_point() + 
  theme_minimal() + 
  MetBrewer::scale_fill_met_c('Hiroshige', direction = -1) + 
  ggtitle("Using super_learner on all of the above with negative log loss")
```
