---
title: "Density Estimation"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(nadir)
```

Thinking about applications of the super learner algorithm 
to causal inference, one key application is through modeling the treatment/exposure 
mechanism. 

In many cases, the treatment/exposure is continuous, necessitating the 
estimation of a "generalized" propensity score — generalized in the sense that 
the treatment/exposure is no longer binary but continuous. 

Here we demonstrate the application of `nadir::super_learner()` in estimating
such a generalized propensity score.

We're not going to worry (yet) about having the most realistic examples; the 
point is that we want to see if we can get density estimation working. 

```{r}
# in order to build a weighting based estimator, we might fit a conditional
# density model of our continuous exposure 
data("Boston", package = "MASS")

# suppose our hypothesis of interest is "does crime reduce median home value?" 
# and we want to control for the following potential confounders (all other variables in the Boston dataset) 
# 
# so we want to regress crime on all the other variables in the Boston dataset 
reg_formula <- as.formula(paste0("crim ~ ", paste0(setdiff(colnames(Boston), c('medv', 'crim')), collapse = ' + ')))

# we would want to regress poverty on these variables

lnr_rf_homoskedastic_density <- function(data, formula, ...) { 
  lnr_homoskedastic_density(data, formula, mean_lnr = lnr_rf, ...)
}

lnr_ranger_homoskedastic_density <- function(data, formula, ...) { 
  lnr_homoskedastic_density(data, formula, mean_lnr = lnr_ranger, ...)
}

lnr_lm_homoskedastic_density <- function(data, formula, ...) { 
  lnr_homoskedastic_density(data, formula, mean_lnr = lnr_lm, ...)
}

lnr_earth_homoskedastic_density <- function(data, formula, ...) { 
  lnr_homoskedastic_density(data, formula, mean_lnr = lnr_earth, ...)
}

learned_sl_density_model <- super_learner(
  data = Boston,
  formula = reg_formula,
  learners = list(
    normal = lnr_lm_density,
    # rf = lnr_rf_homoskedastic_density,
    ranger = lnr_ranger_homoskedastic_density,
    earth = lnr_earth_homoskedastic_density,
    lm = lnr_lm_homoskedastic_density
    ),
  determine_super_learner_weights = determine_weights_using_neg_log_loss,
  verbose = TRUE
)

# compare the learners using negative log likelihood loss
compare_learners(learned_sl_density_model, loss_metric = negative_log_loss)

learned_sl_density_model$learner_weights
```


## Let's validate that conditional density works the way it should 

```{r}


lm_density_predict <- lnr_lm_homoskedastic_density(Boston, reg_formula)

f_lm <- function(ys) {
  x <- Boston[1,]
  sapply(ys, function(y) {
    x[['crim']] <- y
    lm_density_predict(x)
  })
}

integrate(f_lm, min(Boston$crim) - sd(Boston$crim), max(Boston$crim) + sd(Boston$crim), subdivisions = 10000)


earth_density_predict <- lnr_earth_homoskedastic_density(Boston, reg_formula, density_args = list(bw = 30))

f_earth <- function(ys) {
  x <- Boston[1,]
  sapply(ys, function(y) {
    x[['crim']] <- y
    earth_density_predict(x)
  })
}
earth_density_predict(Boston[1,])

integrate(f_earth, min(Boston$crim) - 10*sd(Boston$crim), max(Boston$crim) + 10*sd(Boston$crim))

y_seq <- seq(min(Boston$crim) - 10*sd(Boston$crim), max(Boston$crim) + 10*sd(Boston$crim), length.out = 10000)
# f_earth(y_seq)

delta_y <- y_seq[2]-y_seq[1]

sum(f_earth(y_seq)*delta_y)

```

## Heteroskedastic Learners 

```{r}
#| warning: false
#| message: false
lnr_earth_mean_glm_var_heteroskedastic_density <- function(data, formula, ...) { 
  lnr_heteroskedastic_density(data, formula, mean_lnr = lnr_earth, 
                            var_lnr = lnr_glm, 
                            var_lnr_args = list(family = gaussian(link = 'log')),
                            ...)
}

earth_mean_glm_var_heteroskedastic_predict <- lnr_earth_mean_glm_var_heteroskedastic_density(Boston, reg_formula)

earth_mean_glm_var_heteroskedastic_predict(Boston[1,])

f_earth_glm <- function(ys) {
  x <- Boston[1,]
  sapply(ys, function(y) {
    x[['crim']] <- y
    earth_mean_glm_var_heteroskedastic_predict(x)
  })
}

integrate(f_earth_glm, min(Boston$crim) - 10*sd(Boston$crim), max(Boston$crim) + 10*sd(Boston$crim), subdivisions = 1000)

sum(f_earth_glm(y_seq)*delta_y)

```

```{r}
#| message: false
#| warning: false
lnr_earth_glm_heteroskedastic_density <- function(data, formula, ...) { 
  lnr_heteroskedastic_density(data, formula, mean_lnr = lnr_earth, 
                            var_lnr = lnr_mean, 
                            density_args = list(bw = 3),
                            ...)
}

lnr_earth_glm_heteroskedastic_predict <- lnr_earth_glm_heteroskedastic_density(Boston, reg_formula)

lnr_earth_glm_heteroskedastic_predict(Boston[1,])

f_earth_mean <- function(ys) {
  x <- Boston[1,]
  sapply(ys, function(y) {
    x[['crim']] <- y
    lnr_earth_glm_heteroskedastic_predict(x)
  })
}

# integrate(f_earth, min(Boston$crim) - sd(Boston$crim), max(Boston$crim) + sd(Boston$crim), subdivisions = )

sum(f_earth_mean(y_seq)*delta_y)
```

This indicates that as of now, the `lnr_heteroskedastic_density()` method is not working
correctly.
