---
title: "FAQs"
output: 
  html_document:
    toc: true # table of content true
    toc_depth: 3  # upto three depths of headings (specified by #, ## and ###)
    number_sections: false  ## if you want number sections at each table header
    code_folding: show
    # theme: united  # many options for theme, this one is my favorite.
    # highlight: tango  # specifies the syntax highlighting style
    # css: my.css   # you can add your custom css, should be in same folder
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

We try to cover some anticipated frequently asked questions here. 

```{r setup}
#| echo: false
library(nadir)
```


#### I want to use `super_learner()` for count or nonnegative outcomes.

In principle, you can use `super_learner()` for whatever type of outcomes you want 
as long as a few things hold: 

  - The learners that you pass to `super_learner()` predict that type of outcome. 
  - The loss function used inside the `determine_super_learner_weights()` function
  argument is consistent with the loss function that should be used with your type
  of data. What loss function "should" be used depends on the context, but 
  using the mean-squared-error loss for continuous outcomes and 
  negative log loss for binary outcomes and conditional density models
  is what has been written so far. Refer to the source `R/determine_weights.R`.
  - If using `nadir::super_learner()` for applications in the context of Targeted Learning, 
  it may be useful to understand better the arguments of  
  [*Unified Cross-Validation Methodology For Selection Among Estimators and a General Cross-Validated Adaptive Epsilon-Net Estimator: Finite Sample Oracle Inequalities and Examples* by Mark van der Laan and Sandrine Dudoit, 2003](https://biostats.bepress.com/ucbbiostat/paper130/) first to understand 
  how an appropriate loss function should be chosen depending on the outcome type. 
  My understanding is that some people have used Poisson distribution motivated
  loss functions (<https://discuss.pytorch.org/t/poisson-loss-function/44301/6>, <https://pytorch.org/docs/stable/generated/torch.nn.PoissonNLLLoss.html>) but I 
  would need to think more about if this is the right thing to do in general
  for count outcome data. 
  
#### What are the limitations of `nadir::super_learner()`? 

There are a few key limitations of the design. 

  * Because `learners` (see `?learners`) are understood to be functions that take `data`,
  a `formula` and return a prediction function, there is little to no ability
  (outside of manually following along with the internals of a learner) to 
  check on the internals of learner fits. 
    * That is to say, if you want to peek into the beta coefficients or
    other fit statistics of a learner, this is not supported in
    `nadir::super_learner()` by design.  The reasoning is that an explicit goal
    of `{nadir}` is to keep learner objects lightweight so that building a `super_learner()`
    can be fast.
  * So far, no thought has been put into complex left-hand-sides of regression
  equations. There is no support for left-hand-sides that are not just the 
  name of a column in the `data` passed. The advice for now if you want to 
  model some transformation of the `Y` variable is to apply the transformation
  and store that in the `data` with a column name and to use that new column 
  name in your regression formula(s). 
    * As an explicit subpoint to call attention to, this means so far, no work has been put into supporting survival type outcomes. 
  * So far, everything in `{nadir}` assumes completeness (no missingness) of the data.

#### What if the learner that I want to write really isn't formula based? 

A solution for such a case is to more-or-less ditch the formula piece of a
learner entirely, just treating it as an unused argument, and for your custom
needs you can always build learners that encode details of the structure of your
data.

```{r, class.source='fold-hide'}
data <- matrix(data = rnorm(n = 200), nrow = 20)
colnames(data) <- paste0("X", 1:10)
data <- cbind(data, data %*% rnorm(10))
colnames(data)[ncol(data)] <- 'Y'

lnr_nonformula1 <- function(data, formula, ...) {
  
  # notice by way of knowing things about our data structure, we never reference
  # the formula;  so if you truly don't want to use it, you don't have to.
  
  # as an example, here we do OLS assuming inputs are numeric matrices — 
  # this might even be computationally more performant given how much extra
  # stuff is inside an lm or glm fit.
  X <- as.matrix(data[,grepl(pattern = "^X", colnames(data))])
  Y <- as.matrix(data[,'Y'])
  model_betas <- solve(t(X) %*% X) %*% t(X) %*% Y
  
  learned_predictor <- function(newdata) {
    if ('Y' %in% colnames(newdata)) {
      index_of_y <- which(colnames(newdata) == 'Y')[[1]]
    } else {
      index_of_y <- NULL
    }
    if (is.data.frame(newdata)) {
      newdata <- as.matrix(newdata)
    }
    as.vector(t(model_betas) %*% t(newdata[,-index_of_y, drop=FALSE]))
  }
  return(learned_predictor)
}
attr(lnr_nonformula1, 'sl_lnr_type') <- 'continuous'

# this is essentially a re-implementation of lnr_mean with no reference to the formula 
lnr_nonformula2 <- function(data, formula, ...) {
  
  Y <- data[,'Y']
  Y_mean <- mean(Y)
  
  learned_predictor <- function(newdata) {
    rep(Y_mean, nrow(newdata))
  }
  return(learned_predictor)
}
attr(lnr_nonformula2, 'sl_lnr_type') <- 'continuous'

learned_super_learner <- super_learner(
  data = data,
  learners = list(
    nonformula1 = lnr_nonformula1,
    nonformula2 = lnr_nonformula2),
  formulas = . ~ ., # it doesn't matter what we put here, because neither 
  # learner uses their formula inputs. 
  y_variable = 'Y',
  verbose = TRUE
  )

# observe that the OLS model gets all the weight because it's the correct model:
round(learned_super_learner$learner_weights, 10) 


rm(data) # cleanup
```

So you can see from the immediately prior code snippet, 
if you have some niche application where you would like to avoid 
using the `formulas` argument to `nadir::super_learner()` at all, you can do this
by taking advantage of what you know about how you're going to structure the `data` 
argument.

#### Can I use `{origami}` with `{nadir}`? 

Yes, you can. 

There's a wrapper provided for working with the `folds_*` functions from `{origami}`. 

The first example below is a bit boring, but it does internally use `origami::folds_vfold`. 
The second example demonstrates how to pass another `fold_*` function from the `{origami}` 
package, and any extra arguments passed to `cv_origami_schema` get passed on to the 
`origami::folds_*` function. 

```{r}
#| class-source: 'fold-show'
sl_model <- super_learner(
  data = mtcars,
  formula = mpg ~ cyl + hp,
  learners = list(rf = lnr_rf, lm = lnr_lm, mean = lnr_mean),
  cv_schema = cv_origami_schema,
  verbose = TRUE
)

# if you want to use a different origami::folds_* function, pass it into cv_origami_schema
sl_model <- super_learner(
  data = mtcars,
  formula = mpg ~ cyl + hp,
  learners = list(rf = lnr_rf, lm = lnr_lm, mean = lnr_mean),
  cv_schema = \(data, n_folds) {
    cv_origami_schema(data, n_folds, fold_fun = origami::folds_loo)
  },
  verbose = TRUE
)
```

#### Are there potentially 'sharp edges' to `{nadir}` worth knowing about? 

Yes! Though `{nadir}` tries to make the process user-friendly, there may be 
unexpected behaviors if you use it outside its design-scope and tested 
functionality. 

For example, `{nadir}` and its learners are so far only built to
handle regression formulas where the left-hand-side appears as a column name of
the data passed in. That means transformations of the outcome variable implied
through the formula are not supported. Our advice is that if you want to
transform the outcome variable, you should store that in a column in your data
and run `super_learner()` using that column name in your formula(s).

Another sharp-edge is around the meta-learning step. For example, 
predicting continuous outcomes, one should specify to `super_learner()` that
`outcome_type = 'continuous'` (the default) so that non-negative least
squares is used to minimize a linear combination of the candidate learners
based on held-out mean squared error for the loss function. Additionally, by 
setting `outcome_type = 'binary'` or `outcome_type = 'density'` the 
negative log likelihood / negative log predicted density are used respectively as loss functions. 
In each of these cases, these defaults translate to setting the 
`determine_weights_for_super_learner` function argument appropriately
to one of `determine_super_learner_weights_nnls()`, `determine_weights_for_binary_outcomes`,
`determine_weights_using_neg_log_loss`. These loss functions are 
selected based on work in the loss based estimation literature, especially[^1] [^2] [^3]. 


[^1]: Dudoit, S., & van der Laan, M. J. (2005). Asymptotics of cross-validated risk estimation in estimator selection and performance assessment. In Statistical Methodology (Vol. 2, Issue 2, pp. 131–154). Elsevier BV. https://doi.org/10.1016/j.stamet.2005.02.003  

[^2]: van der Laan, Mark J.; Dudoit, Sandrine; and Keles, Sunduz, "Asymptotic Optimality of Likelihood Based Cross-Validation" (February 2003). U.C. Berkeley Division of Biostatistics Working Paper Series. Working Paper 125.
https://biostats.bepress.com/ucbbiostat/paper125

[^3]: van der Laan, Mark J. and Dudoit, Sandrine, "Unified Cross-Validation Methodology For Selection Among Estimators and a General Cross-Validated Adaptive Epsilon-Net Estimator: Finite Sample Oracle Inequalities and Examples" (November 2003). U.C. Berkeley Division of Biostatistics Working Paper Series. Working Paper 130.
https://biostats.bepress.com/ucbbiostat/paper130


