---
title: "Glmnets and Multipredictors"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`glmnet` (<https://glmnet.stanford.edu/articles/glmnet.html>) and the Highly Adaptive Lasso (<https://tlverse.org/hal9001/>) represent examples of an interesting class of 
statistical learners which have the default behavior of automatically specifying
a grid of tuning parameters `lambda`, fitting models along that grid of `lambda` values,
and optionally-but-often using a CV-optimal selected model from that grid. 

To deal with learners like this in `{nadir}`, we have provided a 3-fold strategy for `{glmnet}`.

Strategy 1: Simplify things a bit  

  * `lnr_glmnet(data, formula, lambda = 0.1)` is very simple and a bit dumb, supporting only one `lambda` value

Strategy 2: Implement what the people want!

  * `lnr_cvglmnet(data, formula, ...)` does what users may expect, which is for each fitting call to 
  result in `{glmnet}` automatically setting the lambda grid (if the 
  lambda argument is not set by the user) and using the CV-optimal selected one to predict. 
  That is to say, `s = 'lambda.min'` is a default setting for the returned predictor from calling
  `lnr_cvglmnet(data, formula)`. 
  
Strategy 3: The pedantic among us may want something else... 

  * If what you think is that `super_learner()` is already doing cross-validation, so 
why should `glmnet::glmnet()` or `hal9001::fit_hal()` be doing another layer of cross-validation,
then we have a solution for you: you specify the `lambda` grid and use `lnr_glmnet_grid()` 
which returns a `nadir_multi_predictor` object. `nadir_multi_predictor` objects are learners that return a list of prediction functions indexed by the lambda grid (vector) which are
automatically expanded inside `nadir::super_learner()` (and `nadir::crossfit_super_learner()`, etc.). 
This way, if your grid is say `lambda = 10^c(0, -1, -2, -3)`, `super_learner()` can 
expressly choose among different `glmnet`s by `lambda` value. 

  * So you can use `lnr_glmnet_grid(data, formula, lambda = c(...)), ...)` too.  


For the Highly Adaptive Lasso, things are similar, but slightly simpler due to 
how `hal9001::fit_hal()` behaves. 

  * `lnr_hal(data, formula, ...)` is supports both one `lambda` value or being left as
  its default `NULL` and automatically selecting a CV-optimal `lambda`. 
  * `lnr_hal_grid(data, formula, ...)` acts similarly to `lnr_glmnet_grid(data, formula, lambda = c(...)))`
  
## Here are some `{glmnet}` examples to make things concrete 

Say you want to super learn with some `glmnet` learners on a grid of lambdas. 

Then you may want `lnr_cvglmnet()` -- it's quite easy to use. 

```{r}
library(nadir)
set.seed(1234)

df = data.frame(Y = rbinom(100, 1, 0.5), 
                A = rbinom(100, 1, 0.5), 
                L1 = runif(100), L2 = runif(100))
sl_fit_a <- super_learner(
  data = df,
  formulas = Y ~ A + L1 + L2,
  outcome_type = "binary",
  n_folds = 10,
  learners = list(
    lnr_cvglmnet, # the object of our present discussion
    lnr_glm,
    lnr_mean
  ),
  extra_learner_args = list(
    list(alpha = 0, lambda = NULL, nlambda = 100) # you don't have to specify lambda --
    # a grid will be made for you by the `{glmnet}` package. 
    # we give nlambda as a hyperparameter as an example.
  )
)
round(sl_fit_a$learner_weights, digits = 4)
```

A downside of this approach is that you can't really see anything
about what lambda the `glmnet::cv.glmnet()` call picked out.

Just to show you, there are some arguments you can pass to the predictor
beyond `newdata` here, just in case you need them.  You can change the `s = 'lambda.min'` approach, but I 
don't know when that would be recommended. Inside `{glmnet}`, I believe the 
default is to use `s = 'lambda.1se'`, which is to predict under the
largest lambda value that is estimated to be within 1 standard deviation
of the CV selected lambda. Say you wanted to do that for some reason. 
Then you would write something like this: 

```{r}
library(tibble)
df <- tibble(
  L1 = rnorm(n = 100), L2 = rnorm(n = 100), A = rnorm(n = 100, mean = L2), Y = 
    rbinom(n = 100, size = 1, plogis(0.1 * L2 + 1.5*A)))

cvglmnet_pred <- lnr_cvglmnet(df, Y ~ A + L1 + L2, family = binomial(link = 'logit')) 
head(cvglmnet_pred(df, s = 'lambda.1se', type = 'response')) 
# you can check that this is not the same as our default, which is to use lambda.min,
# the CV selected lambda
head(cvglmnet_pred(df, type = 'response')) 
head(cvglmnet_pred(df, s = 'lambda.min', type = 'response')) 
```

---

Suppose you want to see what lambda values are selected then, along the lines of 
our Strategy 3. 

Then consider using the lambda grid approach.
This time, we'll make the data generating process a little more `glmnet`
friendly.

```{r}
library(tibble)

n <- 300
df <- tibble(
  L1 = rnorm(n = n),
  L2 = rnorm(n = n),
  L3 = rnorm(n = n),
  L4 = rnorm(n = n),
  L5 = rnorm(n = n),
  A = rbinom(n = n, size = 1, prob = plogis(L4)),
  Y = rbinom(n = n, size = 1, prob = plogis(0.25 * L4 + 0.5 * L5 + 1.25 * A)))

sl_fit_b <- super_learner(
  data = df,
  formulas = Y ~ A + L1 + L2 + L3 + L4 + L5,
  outcome_type = "binary",
  n_folds = 10,
  learners = list(
    lnr_glmnet_grid, # the object of our present discussion
    lnr_glm,
    lnr_mean
  ),
  extra_learner_args = list(
    glmnet_grid = list(lambda = 10^{seq(0, 1, l = 10) * -4}) 
    # you do have to specify lambda now
  )
)
learner_weights <- sl_fit_b$learner_weights
t(t(round(sl_fit_b$learner_weights, digits = 4))) # rotate into a matrix with rownames for viewability
```

In that last output, you can see evidence of how the `nadir_multi_predictor` works.

If you really wanted to dig into one, you could take a look at what 
`lnr_glmnet_grid()` returns:

```{r}
# the output of calling this learner on data with a formula is a "multipredictor" 
# i.e., a list of predictors
multipreds <- lnr_glmnet_grid(
    data = df, 
    formula = Y ~ A + L1 + L2 + L3 + L4 + L5,
    lambda =  10^{seq(0, 1, l = 10) * -4},
    family = binomial(link = 'logit'))

# check out the multipredictors -- they're a list of predictors
names(multipreds)

# make predictions from the learner associated with the 5th lambda from the grid
head(multipreds[[5]](df))
# these are what `super_learner()` sees and cross-validates over
```

## Let's do some Highly Adaptive Lasso examples

```{r}
# using the same df from above;  lnr_hal_grid also returns multipredictors
multipreds <- lnr_hal_grid(
    data = df, 
    formula = Y ~ A + L1 + L2 + L3 + L4 + L5,
    lambda =  10^{seq(0, 1, l = 10) * -4},
    family = binomial(link = 'logit'))

# so we can super learn with it;  though slower. 
timing_c_start <- Sys.time()
sl_fit_c <- super_learner(
  data = df,
  formulas = Y ~ A + L1 + L2 + L3 + L4 + L5,
  outcome_type = "binary",
  n_folds = 3, # reduce folds so it isn't so slow
  learners = list(
    lnr_hal_grid, # the object of our present discussion
    lnr_glm,
    lnr_mean
  ),
  extra_learner_args = list(
    hal_grid = list(lambda = 10^{seq(0, 1, l = 3) * -4}) 
    # user a smaller grid here so it's not so slow
  )
)
timing_c_end <- Sys.time()
print(paste0("timing: "))
print(timing_c_end - timing_c_start)

learner_weights <- sl_fit_c$learner_weights
t(t(round(learner_weights, digits = 3))) # rotate into a matrix with rownames for viewability
```

Or we could have used `lnr_hal()` and let it tune itself.

```{r}
timing_d_start <- Sys.time()
sl_fit_d <- super_learner(
  data = df,
  formulas = Y ~ A + L1 + L2 + L3 + L4 + L5,
  outcome_type = "binary",
  n_folds = 3, # reduce folds so it isn't so slow
  learners = list(
    lnr_hal, # the object of our present discussion
    lnr_glm,
    lnr_mean
  ),
  extra_learner_args = list(
    hal = list(lambda = NULL) # this is the default just made explicit for explanation here 
   )
)
timing_d_end <- Sys.time()
print(paste0("timing: "))
print(timing_c_end - timing_c_start)

learner_weights <- sl_fit_d$learner_weights
t(t(round(learner_weights, digits = 3))) # rotate into a matrix with rownames for viewability
```
