---
title: "Guidance for Developers"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

At present, this is just a handful of thoughts or 
points I thought worth recording about the development of 
`{nadir}`. 

  - I think debugging function factories is going to be hard. 
  How to improve the user experience here? 
  
### On the topic of `{pkgdown}` 

  - From my experience, when I started running into issues compiling the `{pkgdown}` site,
  (errors saying that the .Rd files could not be parsed), I needed to update the dependencies
  for `{pkgdown}`. I ran 
  `install.packages("pkgdown", dependencies = TRUE)` and that fixed the issues I was running
  into towards the start of the `{pkgdown}` setup.. 
  - I think to debug rendering the `{pkgdown}` website, 
  I also ended up using the advice here: <https://stackoverflow.com/questions/66806694/pkgdown-fails-parsing-rd-files-when-examples-are-added>, namely to make sure I have the development version of `downlit` installed.
  
```r
library(devtools)
install_github('r-lib/downlit')
```

  - Because the articles for the `{pkgdown}` website do take some time to compile, 
  I find it useful to know we can run each of the sub-components of `pkgdown::build_site()` 
  individually.  See: <https://pkgdown.r-lib.org/reference/build_site.html>
  - In order to get the `{pkgdown}` website to render the math properly 
  in the documentation for `nadir::lnr_homoskedastic_density()` I had to follow
  the advice in here <https://github.com/r-lib/pkgdown/issues/2704> from user 
  @louisaslett who said that they needed to include in `_pkgdown.yml` manually:

```yaml
template:
  bootstrap: 5
  includes:
    in_header: |
      <link rel="stylesheet" href="https://cdn.jsdelivr.net/npm/katex@0.16.11/dist/katex.min.css" integrity="sha384-nB0miv6/jRmo5UMMR1wu3Gz6NLsoTkbqJghGIsx//Rlm+ZU03BU6SQNC66uf4l5+" crossorigin="anonymous">
      <script defer src="https://cdn.jsdelivr.net/npm/katex@0.16.11/dist/katex.min.js" integrity="sha384-7zkQWkzuo3B5mTepMUcHkMB5jZaolc2xDwL6VFqjFALcbeS9Ggm/Yr2r3Dy4lfFg" crossorigin="anonymous"></script>
      <script defer src="https://cdn.jsdelivr.net/npm/katex@0.16.11/dist/contrib/auto-render.min.js" integrity="sha384-43gviWU0YVjaDtb/GhzOouOXtZMP/7XUzwPTstBeZFe/+rCMvRwr4yROQP43s0Xk" crossorigin="anonymous" onload="renderMathInElement(document.body);"></script>
```

  - For now, the images in the README.Rmd are just embedded manually after
  having been generated (Feb 27 2025) based on the lack of support for 
  Rmd generated images getting copied to the site docs by pkgdown discussed
  in this thread <https://github.com/r-lib/pkgdown/issues/133> with 
  the advice being to put images in `man/figures/` so that images can 
  also be rendered on CRAN.
  - I have tried following the advice here <https://github.com/r-lib/pkgdown/issues/995> to 
  get the articles to be in better ordering, but it didn't seem to work for me. 

```{r setup}
# library(nadir)
```
