---
title: "Timing Learners via the Warning System"
editor_options: 
  chunk_output_type: inline
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`nadir::super_learner()` never lets learner warnings hit your console: every
warning signaled during training or prediction is muffled, captured, and
returned to you in `$warnings_from_training_cv_stage1`,
`$warnings_from_predicting_cv_stage2`, and
`$warnings_from_training_on_entire_data`, each a list of warning *condition
objects* named by the learner that signaled them.

A fun consequence: since warnings are captured as full condition objects (not
just messages), the warning system doubles as a **structured side-channel out
of the training stage**. Anything a learner wants to report about its own
fitting process — without disturbing its return value, which must remain a
prediction function — can be smuggled out as a classed warning condition
carrying extra fields.

In this vignette we use that side-channel to answer a practical question:
*which of my candidate learners are the slowest to fit?* We'll write a
learner-modifier, `add_learner_timing_warnings()`, in the same spirit as
`nadir::add_screener()`: it takes a learner and returns a new learner that
behaves identically, except that it times its training step and signals a
formatted warning carrying the timing results.

```{r setup, message = FALSE}
library(nadir)
library(ggplot2)
```

## Building `add_learner_timing_warnings()`

First, a small constructor for our custom warning condition. Giving the
condition its own class, `learner_timing_warning`, is what lets us later
distinguish *deliberate* timing warnings from any organic warnings the
learners happen to throw (rank-deficiency complaints, convergence grumbles,
and so on). Alongside the human-readable `message`, we stash the raw
`elapsed` seconds and the learner's name as structured fields, so no
message-parsing is needed downstream.

```{r}
learner_timing_warning <- function(learner_name, timing) {
  structure(
    class = c("learner_timing_warning", "warning", "condition"),
    list(
      message = sprintf(
        "[learner timing] %s: trained in %.3f seconds (elapsed) on %s",
        learner_name,
        timing[["elapsed"]],
        format(Sys.time(), "%H:%M:%OS2")
      ),
      call = sys.call(-1),
      learner_name = learner_name,
      elapsed = unname(timing[["elapsed"]])
    )
  )
}
```

Now the modifier itself. Like all `{nadir}` learners, the wrapped version is a
function factory taking `(data, formula, ...)` and returning a prediction
function; the only change is a `system.time()` around the training call and a
`warning()` signaling our classed condition afterwards. We copy over the
`sl_lnr_name` and `sl_lnr_type` attributes so that automatic learner naming
and outcome-type checking keep working exactly as they would for the unwrapped
learner.

```{r}
add_learner_timing_warnings <- function(learner) {
  learner_name <- attr(learner, 'sl_lnr_name')
  if (is.null(learner_name)) {
    learner_name <- 'unnamed_lnr'
  }

  timed_learner <- function(data, formula, ...) {
    timing <- system.time({
      trained_predictor <- learner(data = data, formula = formula, ...)
    })
    warning(learner_timing_warning(learner_name, timing))
    return(trained_predictor)
  }

  attr(timed_learner, 'sl_lnr_name') <- learner_name
  attr(timed_learner, 'sl_lnr_type') <- attr(learner, 'sl_lnr_type')
  return(timed_learner)
}
```

Note that because `super_learner()` muffles the warnings it captures, wrapping
learners this way adds *no console noise at all* — even under parallel
`{future}` plans, where the capture happens inside the workers before results
are collected.

## Fitting a super learner on the `penguins` data

We'll use the complete cases of the `palmerpenguins` data and a library of six
learners of rather different computational temperaments, from the essentially
free `lnr_mean` up through random forests and gradient boosting.

```{r}
penguins <- palmerpenguins::penguins
penguins <- penguins[complete.cases(penguins), ]

flipper_length_formula <-
  flipper_length_mm ~ species + island + bill_length_mm +
    bill_depth_mm + body_mass_g + sex

timed_learners <- lapply(
  list(lnr_mean, lnr_lm, lnr_rf, lnr_earth, lnr_cvglmnet, lnr_xgboost,
       lnr_svm, lnr_bart, lnr_lightgbm, lnr_ranger,
       lnr_earth),
  add_learner_timing_warnings)

set.seed(2026)
sl_model <- super_learner(
  data = penguins,
  formula = flipper_length_formula,
  learners = timed_learners
)
```

Nothing was printed — but the timing warnings were captured. With the default
`n_folds = 5`, each learner signals five warnings during the cross-validation
training stage (one per training fold), plus one more when it is refit on the
entire dataset:

```{r}
length(sl_model$warnings_from_training_cv_stage1)
length(sl_model$warnings_from_training_on_entire_data)

# the warnings are named by learner, and the messages are human-readable:
conditionMessage(sl_model$warnings_from_training_cv_stage1[[1]])
```

## Extracting the timing warnings

Now we harvest our deliberate warnings back out. We filter each warning field
for conditions inheriting from `learner_timing_warning` — leaving any organic
learner warnings untouched — and assemble the structured `$elapsed` fields
into a tidy data frame.

```{r}
extract_learner_timings <- function(sl_model) {
  warning_fields <- c(
    cv_training = 'warnings_from_training_cv_stage1',
    full_data_fit = 'warnings_from_training_on_entire_data')

  do.call(rbind, lapply(names(warning_fields), function(stage) {
    warnings_list <- sl_model[[warning_fields[[stage]]]]
    timing_warnings <- Filter(
      function(w) inherits(w, 'learner_timing_warning'),
      warnings_list)
    if (length(timing_warnings) == 0) return(NULL)
    data.frame(
      learner = vapply(timing_warnings, `[[`, character(1), 'learner_name'),
      stage = stage,
      elapsed = vapply(timing_warnings, `[[`, numeric(1), 'elapsed')
    )
  }))
}

learner_timings <- extract_learner_timings(sl_model)
head(learner_timings)
```

## Visualizing which learners are fastest to fit

Averaging the per-fold training times gives us a bar chart of mean
time-to-fit per learner, with the individual fold timings overlaid as points
so the fold-to-fold variability is visible too.

```{r, fig.width = 8, fig.height = 4.5}
cv_timings <- learner_timings[learner_timings$stage == 'cv_training', ]

mean_timings <- aggregate(elapsed ~ learner, data = cv_timings, FUN = mean)
mean_timings$learner <- reorder(mean_timings$learner, mean_timings$elapsed)
cv_timings$learner <- factor(cv_timings$learner, levels = levels(mean_timings$learner))

ggplot(mean_timings, aes(x = elapsed, y = learner, fill = learner)) +
  geom_col(alpha = .8) +
  geom_point(data = cv_timings, shape = 21, fill = 'white', alpha = .75) +
  # scale_fill_brewer(palette = 'Set2', guide = 'none') +
  scale_fill_brewer(palette = 'BrBG', guide = 'none') +
  
  theme_bw() +
  labs(
    x = "mean elapsed training time per fold (seconds)",
    y = NULL,
    title = "Time-to-fit for each candidate learner",
    subtitle = "Bars show the mean across the 5 CV training folds; points show individual folds",
    caption = "Task: flipper_length_mm ~ . on the penguins complete cases, 333 rows.
Timings harvested from deliberately-signaled learner_timing_warning conditions
captured by nadir::super_learner()'s warning system.") +
  theme(plot.caption.position = 'plot')
```

Unsurprisingly, `lnr_mean` and `lnr_lm` are effectively free, while the
tree-ensemble learners dominate the fitting budget. Since timing warnings are
signaled once per training call, this approach extends without modification to
`cv_super_learner()` and `crossfit_super_learner()`, whose outputs aggregate
the same captured-warning fields across their outer folds.

Two closing remarks on the pattern itself. First, a warning is not an error:
a timing-wrapped learner keeps its weight-eligibility, its predictions, and
its place in the ensemble entirely unchanged — we can confirm the wrapped
learners were named and weighted as usual:

```{r}
round(sl_model$learner_weights, 3)
```

Second, nothing about this trick is specific to timing. Any per-fit
diagnostic — memory use, number of boosting rounds actually run, screening
decisions — can ride the same classed-condition side-channel out of the
training stage and back to you through `super_learner()`'s captured warnings.
