---
title: "Translation Guide: Coming from `{SuperLearner}` or `{sl3}`"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

This article is the translation guide for users arriving at `{nadir}` from the
[`{SuperLearner}`](https://cran.r-project.org/package=SuperLearner) or
[`{sl3}`](https://tlverse.org/sl3/) packages. All three packages implement
the same algorithm — the super learner of van der Laan, Polley & Hubbard
(2007), so hopefully users' statistical knowledge transfers directly and easily
with only the interface differing between them.

The code chunks in this article are shown side-by-side for reference and are
not evaluated here. For evaluated head-to-head comparisons, including
checks that the three packages produce closely agreeing results, see the
[Comparison to SuperLearner and sl3](comparison_to_SuperLearner.html) and
[Benchmarking](Benchmarking.html) articles.

If any terminology below is unfamiliar, see the [Glossary](Glossary.html).

# The Big Picture: Three Interface Philosophies

The three packages wrap the same algorithm in three different interface
styles:

* **`{SuperLearner}`** separates the outcome and predictors into a `Y`
  vector and an `X` data.frame, names learners as strings
  (`"SL.glmnet"`), and configures behavior through `family`, `method`, and
  `cvControl` arguments.
* **`{sl3}`** is object-oriented (R6): you construct a `Task` object
  bundling the data with metadata, instantiate learner objects
  (`Lrnr_glmnet$new()`), compose them into a `Stack`, wrap the stack in
  `Lrnr_sl` with a metalearner, and call `$train()`.
* **`{nadir}`** is functional: `super_learner()` takes the `data` directly,
  a `formula` (or a *list of formulas, one per learner*), and a list of
  learner *functions* (`lnr_glmnet`). Learners are plain functions that take
  `(data, formula, ...)` and return a prediction function — no strings, no
  R6 classes.

The practical consequences of `{nadir}`'s formula-first design:

* **No design matrix construction.** You never split your data into `Y` and
  `X`. Setting variables as factors, transformations, and interactions are easily 
  expressed in the formula as in the style of `glm()` and other formula-based 
  statistical learning packages in R.
* **Different learners can get different formulas**, which is what makes it
  natural to include `lme4`-style random effects (`(1 | group)`) or
  `mgcv`-style smooths (`s(x1, x2)`) as candidate learners -- something that
  has remained difficult in both `{SuperLearner}` and `{sl3}`.
* **Writing a custom learner is a few lines of code** (a function factory),
  rather than an `SL.*`/`predict.SL.*` function pair or an R6 subclass.

# Side-by-Side: Fitting a Super Learner

Here is the same task — regress `medv` on all other variables in the Boston
housing data with a mean, glmnet, and ranger library — in all three packages.

### `{SuperLearner}`

```{r}
library(SuperLearner)
data(Boston, package = "MASS")

sl <- SuperLearner(
  Y = Boston$medv,
  X = subset(Boston, select = -medv),
  SL.library = c("SL.mean", "SL.glmnet", "SL.ranger"),
  cvControl = list(V = 10)
)
```

### `{sl3}`

```{r}
library(sl3)
data(Boston, package = "MASS")

task <- make_sl3_Task(
  data = Boston,
  covariates = setdiff(colnames(Boston), "medv"),
  outcome = "medv"
)

stack <- Stack$new(Lrnr_mean$new(), Lrnr_glmnet$new(), Lrnr_ranger$new())
sl <- Lrnr_sl$new(learners = stack, metalearner = Lrnr_nnls$new(),
                  cv_control = list(V = 10))
sl_fit <- sl$train(task)
```

### `{nadir}`

```{r}
library(nadir)
data(Boston, package = "MASS")

sl_fit <- super_learner(
  data = Boston,
  formulas = medv ~ .,
  learners = list(mean = lnr_mean, glmnet = lnr_glmnet, ranger = lnr_ranger),
  n_folds = 10
)
```


# Concept-by-Concept Translation

| Concept | `{SuperLearner}` | `{sl3}` | `{nadir}` |
|---|---|---|---|
| Fit a super learner | `SuperLearner(Y, X, SL.library, ...)` | `Lrnr_sl$new(...)$train(task)` | `super_learner(data, formulas, learners, ...)` |
| Specify the data | `Y` vector + `X` data.frame | `make_sl3_Task(data, covariates, outcome)` | `data` + `formulas` |
| Specify the library | character vector `SL.library` | `Stack$new(Lrnr_a$new(), ...)` | `learners = list(a = lnr_a, ...)` |
| A single learner | `"SL.glmnet"` (string) | `Lrnr_glmnet$new()` (R6 object) | `lnr_glmnet` (function) |
| Outcome family/type | `family = gaussian()` / `binomial()` | `outcome_type` on the task | `outcome_type = 'continuous'` / `'binary'` / `'multiclass'` / `'density'` |
| Number of CV folds | `cvControl = list(V = 10)` | `cv_control = list(V = ...)` / `folds` on the task | `n_folds` argument (default: 5) |
| Meta-learning method | `method = "method.NNLS"` etc. | `metalearner = Lrnr_nnls$new()` etc. | `determine_super_learner_weights = ...` (defaults chosen by `outcome_type`) |
| Learner hyperparameters | `create.Learner()` / writing wrappers | params to `Lrnr_*$new(...)` | `extra_learner_args = list(...)` |
| Observation weights | `obsWeights = w` | `weights` column declared on the task | `weights` argument passed through to all learners |
| Clustered/dependent data | `id = cluster_id` | `id` on the task / custom `folds` | `cluster_ids = ...` |
| Stratified CV | `cvControl = list(stratifyCV = TRUE)` (binary Y) | custom `folds` via `{origami}` | `strata_ids = ...` |
| Custom CV structure | `cvControl = list(validRows = ...)` | `folds = origami::make_folds(...)` on the task | `cv_schema = ...` (e.g. `cv_origami_schema`) |
| Screening | `SL.library = list(c("SL.glmnet", "screen.corP"))` | `Pipeline$new(Lrnr_screener_*, lrnr)` | `add_screener(lnr_glmnet, screener_cor)` |
| Predict on new data | `predict(sl, newdata = ...)` | `sl_fit$predict(new_task)` | `sl_fit$predict(newdata)` or `predict(sl_fit, newdata)` |
| Ensemble weights | `sl$coef` | `sl_fit$coefficients` | `sl_fit$learner_weights` or `coef(sl_fit)` |
| Per-learner CV risk | `sl$cvRisk` | `sl_fit$cv_risk(eval_fun)` | `compare_learners(sl_fit)` |
| CV the whole super learner | `CV.SuperLearner(...)` | `cv_sl(lrnr_sl = sl_fit, eval_fun = ...)` | `cv_super_learner(...)` |
| Discrete super learner | reported by `CV.SuperLearner` | `Lrnr_cv_selector` as metalearner | `ensemble_or_discrete = 'discrete'` |
| Parallelization | `mcSuperLearner()` / `snowSuperLearner()` / `parallel =` | `{future}` plans | `{future}` plans |

# Learner Name Translation

The table below maps commonly used learners. `{nadir}` learner functions are
documented at `?learners`, `?binary_learners`, `?multiclass_learners`, and
`?density_learners`.

| Model | `{SuperLearner}` | `{sl3}` | `{nadir}` |
|---|---|---|---|
| Outcome mean | `SL.mean` | `Lrnr_mean` | `lnr_mean` |
| Linear model | `SL.lm` | `Lrnr_glm` | `lnr_lm` |
| GLM | `SL.glm` | `Lrnr_glm` | `lnr_glm` |
| Logistic regression | `SL.glm` + `family = binomial()` | `Lrnr_glm` (binomial task) | `lnr_logistic` |
| Elastic net (CV-selected lambda) | `SL.glmnet` | `Lrnr_glmnet` | `lnr_cvglmnet` |
| Elastic net (fixed lambda) | — | `Lrnr_glmnet$new(lambda = ...)` | `lnr_glmnet` (+ `lambda` via `extra_learner_args`) |
| Elastic net (every lambda on the path as its own candidate) | — | — | `lnr_glmnet_grid` |
| Random forest (randomForest) | `SL.randomForest` | `Lrnr_randomForest` | `lnr_rf` (binary: `lnr_rf_binary`) |
| Random forest (ranger) | `SL.ranger` | `Lrnr_ranger` | `lnr_ranger` (binary: `lnr_ranger_binary`) |
| MARS (earth) | `SL.earth` | `Lrnr_earth` | `lnr_earth` |
| GAM | `SL.gam` (via `{gam}`) | `Lrnr_gam` (via `{mgcv}`) | `lnr_gam` (via `{mgcv}`, full `s()` syntax in the formula) |
| Gradient boosting (gbm) | `SL.gbm` | `Lrnr_gbm` | `lnr_gbm` |
| XGBoost | `SL.xgboost` | `Lrnr_xgboost` | `lnr_xgboost` |
| LightGBM | — | `Lrnr_lightgbm` | `lnr_lightgbm` |
| Highly adaptive lasso | `SL.hal9001` | `Lrnr_hal9001` | `lnr_hal` (grid: `lnr_hal_grid`) |
| CART | `SL.rpart` | `Lrnr_rpart` | `lnr_rpart` |
| SVM | `SL.svm` | `Lrnr_svm` | `lnr_svm` |
| k-nearest neighbors | `SL.knn` | — | `lnr_knn` |
| BART | `SL.bartMachine` | `Lrnr_dbarts` | `lnr_bart` |
| Neural net (nnet, binary) | `SL.nnet` | `Lrnr_nnet` | `lnr_nnet` |
| Multinomial (multiclass) | — | `Lrnr_multinom`-style learners | `lnr_multinomial_nnet`, `lnr_multinomial_vglm` |
| Mixed models (lme4) | — | — | `lnr_lmer`, `lnr_glmer` |
| Semiparametric conditional density | — | `Lrnr_density_semiparametric` | `lnr_homoskedastic_density`, `lnr_heteroskedastic_density` |

A few rows deserve comment:

* **`SL.glmnet` vs. `lnr_glmnet`.** `SL.glmnet` and `Lrnr_glmnet` internally
  select `lambda` by their own cross-validation (`cv.glmnet`). The direct
  `{nadir}` analogue of that behavior is **`lnr_cvglmnet`**.
  `lnr_glmnet` instead fits a *single, fixed* `lambda` (default `0.2`) — use
  it when you want to expose specific penalty levels as distinct candidates.
  Better still, `lnr_glmnet_grid` fits a whole regularization path in one
  pass and lets the meta-learner weight every `lambda` value separately,
  something neither `{SuperLearner}` nor `{sl3}` exposes directly.
* **`lnr_lmer`, `lnr_glmer`, and `mgcv` smooths** have no counterparts in
  the other packages; supporting these formula-based model families is a
  central motivation for `{nadir}` (see the README's "Why reimplement super
  learner again?").

# Translating Common Workflows

## Binary outcomes

```{r}
# {SuperLearner}
sl <- SuperLearner(Y = mtcars$am, X = mtcars[, c("hp", "wt")],
                   family = binomial(),
                   SL.library = c("SL.mean", "SL.glm"))

# {sl3}: declare outcome_type on the task (or let sl3 infer it)
task <- make_sl3_Task(data = mtcars, covariates = c("hp", "wt"),
                      outcome = "am", outcome_type = "binomial")

# {nadir}
sl_fit <- super_learner(
  data = mtcars,
  formulas = am ~ hp + wt,
  learners = list(mean = lnr_mean, logistic = lnr_logistic),
  outcome_type = 'binary'
)
```

In `{nadir}`, declaring `outcome_type = 'binary'` does three things at once:
it switches the default meta-learning method and loss to negative log loss,
it passes outcome-type-dependent arguments to learners that need them (e.g.
`family = 'binomial'` for `lnr_glmnet`; see the `outcome_type_dependent_args`
learner attribute), and it warns if a supplied learner declares itself
unsuited to binary outcomes. There is no `family =` argument to translate
beyond this.

## Hyperparameter tuning

In `{SuperLearner}`, tuning over hyperparameter values typically goes through
`create.Learner()` or hand-written wrapper functions; in `{sl3}`, you
instantiate the same `Lrnr_*` class several times with different parameters.
In `{nadir}`, you list the same learner function several times under
different names and supply per-learner arguments through
`extra_learner_args`:

```{r}
sl_fit <- super_learner(
  data = Boston,
  formulas = medv ~ .,
  learners = list(
    mean    = lnr_mean,
    glmnet1 = lnr_glmnet,
    glmnet2 = lnr_glmnet,
    rf      = lnr_rf
  ),
  extra_learner_args = list(
    .default = NULL,
    glmnet1  = list(lambda = 0.1),
    glmnet2  = list(lambda = 0.5),
    rf       = list(ntree = 500)
  )
)
```

For penalized regression specifically, prefer `lnr_glmnet_grid` /
`lnr_hal_grid`, which fit the entire path in a single call and expand each
grid value into its own weighted pseudo-learner.

## Screening

```{r}
# {SuperLearner}: pair learners with screeners inside SL.library
sl <- SuperLearner(Y = y, X = x,
                   SL.library = list(c("SL.glmnet", "screen.corP"),
                                     "SL.mean"))

# {sl3}: compose a screener and a learner into a Pipeline
screened_lrnr <- Pipeline$new(Lrnr_screener_correlation$new(), Lrnr_glmnet$new())

# {nadir}: build a screened learner with add_screener()
lnr_glmnet_screened <- add_screener(
  learner = lnr_glmnet,
  screener = screener_cor,
  screener_extra_args = list(threshold = 0.2)
)

sl_fit <- super_learner(
  data = Boston,
  formulas = medv ~ .,
  learners = list(glmnet_screened = lnr_glmnet_screened, mean = lnr_mean)
)
```

See `?screeners` for the available screening functions
(`screener_cor`, `screener_cor_top_n`, `screener_t_test`).

## Clustered data and custom cross-validation

`{SuperLearner}`'s `id` argument and `{sl3}`'s task-level `id` /
`origami`-based `folds` both translate to `{nadir}`'s `cluster_ids`,
`strata_ids`, and `cv_schema` arguments:

```{r}
# {SuperLearner}
sl <- SuperLearner(Y = y, X = x, SL.library = lib, id = df$cluster_id)

# {sl3}
task <- make_sl3_Task(data = df, covariates = covs, outcome = "y",
                      id = "cluster_id")

# {nadir}
sl_fit <- super_learner(
  data = df,
  formulas = y ~ .,
  learners = learners,
  cluster_ids = df$cluster_id
)
```

Behind the scenes `{nadir}` also uses `{origami}` (the same fold machinery
`{sl3}` uses) via `cv_origami_schema`, so `{sl3}` users who relied on
`origami::folds_*` functions (leave-one-out, timeseries folds, etc.) can keep
using them:

```{r}
sl_fit <- super_learner(
  data = df,
  formulas = y ~ .,
  learners = learners,
  cv_schema = \(data, n_folds) {
    cv_origami_schema(data, n_folds, fold_fun = origami::folds_loo)
  }
)
```

See the [Clustered and Dependent Data](Clustered-Data.html) article.

## Predicting, inspecting weights, and comparing learners

```{r}
# {SuperLearner}
preds <- predict(sl, newdata = newdata)$pred
sl$coef     # ensemble weights
sl$cvRisk   # per-learner cross-validated risk

# {sl3}
pred_task <- make_sl3_Task(data = newdata, covariates = covs, outcome = "y")
preds <- sl_fit$predict(pred_task)
sl_fit$coefficients
sl_fit$cv_risk(loss_squared_error)

# {nadir}
preds <- sl_fit$predict(newdata)       # or predict(sl_fit, newdata = newdata)
sl_fit$learner_weights                 # or coef(sl_fit)
compare_learners(sl_fit)               # per-learner held-out loss
```

Two conveniences to note: `{nadir}` does not require constructing a task for
prediction (any data.frame with the needed columns works, and the outcome
column need not be present in `newdata`), and `{nadir}`'s model classes come
with the standard regression S3 methods (`predict()`, `coef()`, `fitted()`,
`residuals()`, `nobs()`, `print()`, `summary()`).

## Cross-validating the super learner itself

```{r}
# {SuperLearner}
cv_sl <- CV.SuperLearner(Y = y, X = x, SL.library = lib, V = 5)
summary(cv_sl)

# {sl3}
cv_fit <- cv_sl(lrnr_sl = sl_fit, eval_fun = loss_squared_error)

# {nadir}
cv_output <- cv_super_learner(
  data = Boston,
  formulas = medv ~ .,
  learners = list(mean = lnr_mean, glmnet = lnr_cvglmnet, ranger = lnr_ranger)
)
cv_output$cv_loss
```

`{nadir}` additionally provides `crossfit_super_learner()` for workflows
(AIPW, TMLE, DML) that require retaining each outer fold's fitted ensemble
and its out-of-fold predictions — see the
[Doubly Robust Estimation](Doubly-Robust-Estimation.html) article.

## Discrete super learner

```{r}
# {nadir}
sl_fit <- super_learner(
  data = Boston,
  formulas = medv ~ .,
  learners = list(mean = lnr_mean, glmnet = lnr_cvglmnet, ranger = lnr_ranger),
  ensemble_or_discrete = 'discrete'
)
```

This gives weight 1 to the learner that won the meta-learning step and weight
0 to all others — the quantity `CV.SuperLearner` reports as the "Discrete SL"
and what `{sl3}` achieves with a CV-selector metalearner.

# Writing Custom Learners: The Biggest Difference

In `{SuperLearner}`, a custom learner is a pair of functions following the
`SL.template` — an `SL.mylearner(Y, X, newX, family, obsWeights, ...)`
fitting function returning `list(pred, fit)`, plus a
`predict.SL.mylearner()` method. In `{sl3}`, a custom learner is an R6 class
inheriting from `Lrnr_base` with private `.train()` and `.predict()`
methods.

In `{nadir}`, a learner is a plain function taking `(data, formula, ...)`
that returns a prediction function of `newdata`:

```{r}
lnr_custom <- function(data, formula, ...) {
  model <- your_fitting_function(formula = formula, data = data, ...)

  function(newdata) {
    predict(model, newdata = newdata)   # a numeric vector, one per row
  }
}

# optional, but recommended:
attr(lnr_custom, 'sl_lnr_name') <- 'custom'
attr(lnr_custom, 'sl_lnr_type') <- 'continuous'
```

That is the complete contract. See the
[Currying, Closures, and Function Factories](currying_closures_and_function_factories.html)
article for why this works and `?learners` for conventions (including
accepting a `weights` argument if the underlying method supports observation
weights).

# Defaults That Differ {#defaults-that-differ}

When porting an analysis, watch for these differences in defaults, which can
produce (legitimately) different numerical results across the packages even
on the same data and library:

* **Number of folds.** `SuperLearner()` defaults to `V = 10`;
  `nadir::super_learner()` defaults to `n_folds = 5`. Set them to match when
  comparing.
* **Fold assignment.** Each package randomizes fold membership differently,
  so results agree in distribution, not to machine precision, even with
  matching seeds.
* **glmnet behavior.** As noted above, translate `SL.glmnet` to
  `lnr_cvglmnet` (not `lnr_glmnet`) to match the internal-CV lambda
  selection.
* **Meta-learner.** All three default to NNLS-type weight determination for
  continuous outcomes, but for binary/multiclass/density outcomes `{nadir}`
  selects a negative-log-loss-based method automatically from
  `outcome_type`; in `{SuperLearner}` the analogous choice is
  `method = "method.NNloglik"` and in `{sl3}` it is set through the
  `metalearner` argument.
* **Missing data.** `{nadir}` deliberately errors on incomplete rows unless
  you opt in via `use_complete_cases = TRUE`, rather than silently handling
  or deleting them.

# See Also

* [Benchmarking](Benchmarking.html) — timing comparisons across dataset
  sizes.
* [Glossary](Glossary.html) — definitions of the technical terms used here.
* `?super_learner`, `?learners`, `?screeners`, `?cv_super_learner`,
  `?crossfit_super_learner`.
