---
title: "Visualizing Performance"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`{nadir}` model objects come with `plot()` methods, so visualizing how your
candidate learners and the super learner ensemble performed takes one line
of code. This article tours the available plot types and explains which
comparison is the *fair* comparison to make.

```{r setup}
library(nadir)
```

```{r, echo=FALSE}
set.seed(1111)
options(future.globals.maxSize = +Inf)
```

Throughout we'll use the Boston housing data and a moderately diverse
library of learners:

```{r}
#| message: false
data("Boston", package = "MASS")

learners <- list(
  lm = lnr_lm,
  rf = lnr_rf,
  ranger = lnr_ranger,
  xgboost = lnr_xgboost,
  glmnet = lnr_glmnet,
  earth = lnr_earth
)

sl_model <- super_learner(
  data = Boston,
  formulas = medv ~ .,
  learners = learners
)
```

## Comparing the candidate learners

`compare_learners()` reports each candidate's cross-validated held-out loss
as a table:

```{r}
compare_learners(sl_model)
```

The default `plot()` method shows the same information visually, along
with how variable each learner's performance was across the
cross-validation folds:

```{r}
#| fig-width: 8
#| fig-height: 5
#| fig-dpi: 300
#| fig-retina: 3
plot(sl_model)
```

For each candidate learner, the bar and filled point show the mean held-out
loss across the cross-validation folds, each circle mark shows one fold's
held-out loss, and the range shows plus or minus 1 standard deviation across folds.
Learners are ordered best (lowest mean loss) at the top.

### So where is super learner in this plot?

Without additional work, we can't fairly judge super learner yet 
because we don't have the *super learner's* held-out loss value.

A `super_learner()` fit chooses its ensemble weights
*using* the candidates' holdout predictions, so scoring the ensemble on
those same holdouts would be optimistically biased. The ensemble gets to
see the data that it would be evaluated on. `plot(sl_model)` therefore
compares only the candidates, which were each scored honestly on held out data.

To score the ensemble itself fairly, we need another layer of
cross-validation. That's the purpose of `cv_super_learner()`.

## Honestly evaluating the super learner

```{r}
#| warning: false
#| message: false
cv_results <- cv_super_learner(
  data = Boston,
  formulas = medv ~ .,
  learners = learners
)
```

```{r}
#| fig-width: 8
#| fig-height: 5
#| fig-dpi: 300
#| fig-retina: 3
plot(cv_results)
```

Now the `super_learner` row appears alongside the candidates, and the
comparison is a level playing field: within each outer fold, the candidate
fits and the ensemble weights were all estimated entirely without the
validation data they're evaluated on, and every learner was trained on the
same outer training data. 

## Ensemble weight stability

How consistently does the super learner weight its candidates when the
training data changes? `type = "weights"` shows the ensemble weights
estimated within each outer fold:

```{r}
#| fig-width: 8
#| fig-height: 5
#| fig-dpi: 300
#| fig-retina: 3
plot(cv_results, type = "weights")
```

Tight clusters indicate an ensemble that is stable under resampling. If a
learner's weights vary wildly from fold to fold, that's a sign the data
don't strongly distinguish it from its competitors (or that it's unstable);
missing points would indicate folds where a learner errored and was dropped.

## Predictions vs. observed outcomes

Finally, `type = "fitted"` plots out-of-fold ensemble predictions against
the observed outcomes, with the identity line for reference:

```{r}
#| fig-width: 6
#| fig-height: 5
#| fig-dpi: 300
#| fig-retina: 3
plot(cv_results, type = "fitted")
```

Every prediction here comes from an ensemble trained entirely without that
observation, so systematic deviations from the identity (45-degree) line would reflect 
shortcomings of the overall super learned model fit. The same view is
available for a plain `super_learner()` fit via
`plot(sl_model, type = "fitted")`, using its cross-validated holdout
predictions.

## Customizing the plots

Each `plot()` method returns a `ggplot` object, so the usual `{ggplot2}`
grammar applies if you want to restyle them:

```{r}
#| fig-width: 8
#| fig-height: 5
#| fig-dpi: 300
#| fig-retina: 3
plot(cv_results) +
  ggplot2::scale_fill_brewer(palette = "Set2") +
  ggplot2::theme_minimal()
```

If you need something the plot methods don't offer, the raw materials
are all retained on the fitted objects: per-fold holdout predictions live
in `sl_model$holdout_predictions`, and the per-fold `super_learner()` fits
from the outer cross-validation live in `cv_results$crossfit`.
